### 2022-10-29 responseToReviews/response_to_reviewers.tex
init

\documentclass[12pt]{report}
\usepackage{color,soul}
\usepackage{subcaption,wrapfig,graphicx,booktabs,fancyhdr,amsmath,amsfonts}
\usepackage{cite,bm,amssymb,amsthm,url,multirow,times,enumitem,comment}
\usepackage{mathtools,siunitx,tikz}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage[colorlinks=true, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
\newcommand{\vb}{\boldsymbol}
\newcommand{\vbh}[1]{\hat{\boldsymbol{#1}}}
\newcommand{\vbc}[1]{\check{\boldsymbol{#1}}}
\newcommand{\vbb}[1]{\bar{\boldsymbol{#1}}}
\newcommand{\vbt}[1]{\tilde{\boldsymbol{#1}}}
\newcommand{\vbs}[1]{{\boldsymbol{#1}}^*}
\newcommand{\vbd}[1]{\dot{{\boldsymbol{#1}}}}
\newcommand{\abs}[1]{\left|{#1}\right|}
\newcommand{\by}{\times}
\newcommand{\tr}{\mathsf{T}}
\newcommand{\sfrac}[2]{\textstyle \frac{#1}{#2}}
\newcommand{\ba}{\begin{array}}
\newcommand{\ea}{\end{array}}
\newcommand{\sinc}{\text{sinc}}
\renewcommand{\equiv}{\triangleq}
\newcommand{\cnr}{C/N_0}
\newcommand{\sgn}{\text{sgn}}
\renewcommand{\Re}{\mathbb{R}}
\renewcommand{\Im}{\mathbb{I}}
\newcommand{\E}[1]{\mathbb{E}\left[ #1 \right]}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\DeclareMathOperator*{\argmin}{\arg\!\min}
\DeclareMathOperator*{\argmax}{\arg\!\max}

\topmargin = 0 mm 
\oddsidemargin = -1 mm 
\evensidemargin = -1 mm
\headheight = 0 mm 
\headsep = 8 mm 
\textheight = 220 mm 
\textwidth =170 mm 
\parindent = 0 mm
\parskip = 4 mm

\begin{document}

\section*{Reviewer 1}
\subsection*{General Comments:}

{\textit{This is a very nice, thorough piece of work that develops on the
    authors' previous contributions to GNSS spoofing detection. The paper
    expands on the previously introduced ``sandwich'' detector, which utilises
    both overall power level monitoring and per channel correlation distortion
    monitoring to classify each received signal as belonging to one of the 4
    classes: nominal, multipath, jammed, or spoofed.}}

{\textit{The development is clear and the text is well structured and well
    written and deserves to be published with some largely minor
    modifications.}}

{Thank you for your assessment and helpful comments.}

\subsection*{Suggested Improvements:}

{\textit{One area that could do with some clarification relates to
      Section IV A -
Received Power Level Measurement, in particular Eqn
      (5). In the preceding
development leading up to this section the
      authors have considered a single
signal only, which is both appropriate
      and simplifies the notation. However,
when considering the overall
      power in the receiver, the chosen notation and
description becomes
      somewhat convoluted. I think it would clarify the matter
here to
      introduce the total power as the sum of the power from all
      satellites
(and all signals within the L1 band from all satellites),
      the power from all
interference sources (jammers/spoofers etc) and the
      cross terms (between
satellites, between interferers and between
      satellites and interferers). In
this manner it is not necessary to
      decide whether an interferer should be
classified as part of the
      MAI. With this fuller notation, needed only for the
power level
      measurement, it would be clear where each contribution to the
overall
      measured power comes from.}}

  { This is a good suggestion.  We have adopted it fully in the revised
    Section IV.  The new Eq. (6) expresses the combined signal as the
    summation of all authentic and interference signals, plus noise.  $P_k$ is
    now defined in terms of this combined signal (cf. Eq. (7)), and the
    mean of $P_k$ (in linear units) is expressed as a sum of constituents,
    including cross-terms between authentic and interference signals
    (cf. Eq. (8)).  No cross-terms between interferers themselves is
    necessary in Eq. (8), as interferers are assumed to be mutually
    independent, as now explained in the paragraph just before (8).}

{\textit{In addition, I would argue that Eqn (6) is not strictly
      correct, as the right
hand side represents not the Power Spectral
      Density due to MAI, but rather the
total power due to MAI. Indeed the
      right hand side is the closed form solution
of the so-called spectral
      separation coefficient for BPSK(1) assuming infinite
transmit and
      receive bandwidths. Using Eqn (6) as is would greatly over
estimate the
      contribution of MAI to total received power.}}

  {We agree, and are grateful that you caught this error.  The revised
    expression for $P_k$ (Eqs. (7) and (8)) does not attempt approximation via
    $M_0$.  The earlier approach was equivalent to the current (correct)
    approach when $W_P = 3/(2\tau_c)$, but, as $W_P$ is assumed larger than
    $3/(2\tau_c)$, our former approach exaggerated the contribution of
    multi-access signals to $P_k$.  We have corrected our code and re-run the
    paper's simulations and tests against empirical data.}

{\textit{This leads to my final concern with this section: I don't really
      see how matched
power spoofing can lead to a total power delta of
      approximately 2 dB in a 2 MHz
bandwidth as shown in Fig. 4. By my
      calculation this would require an average
C/No of about 50 dB-Hz from
      approximately 20 satellites simultaneously.
I presume the large delta
      in Fig. 4 is largely due to the RF image energy
introduced by the
      spoofer. My concern then is that this same spoofer has been
used in the
      live data experiments also. Is this justifiable? Is it likely that
such
      a large amount of in-band noise would be generated by a "real world"
      spoofer?
}}

  We have revisited our empirical measurements and our calculations and can
  re-affirm the $\sim$2 dB uptick in the 2-MHz bandwidth at the beginning of
  spoofing in TEXBAT Scenario 3.  We were also surprised at the size of the
  increase.  It can be explained by noting that, for the high-quality antenna
  (Trimble Geodetic Zephyr II) and receiver (a 16-bit National Instruments
  front end) in question, the total authentic power in the 2-MHz bandwidth
  under nominal (clean) conditions was measured to be approximately -142 dBW.
  This is the combined power of $M_s + 1 = 12$ authentic GPS L1 C/A signals
  (plus a small amount of power from GPS L1(P)).

  This is a surprisingly large combined power.  Assuming a noise floor of -204
  dBW/Hz (thermal noise only, no multi-access noise), the total noise power in
  a 2-MHz bandwidth would be $63 - 204 = -141$ dBW.  Thus, the combined power
  of the authentic signals (and mild multipath) for this antenna/receiver
  combination is nearly equal to the total noise power in band.  One can
  verify this by examining the new Fig. 3, which shows the nominal (clean)
  spectrum in dark gray.  Selective integration of the spectrum reveals that
  the power in the center 2 MHz band is indeed twice that of the power in a 2
  MHz band of the noise floor (which one can approximate by centering a 2 MHz
  band at -4 MHz).

    The upshot of this is that, \emph{relative to the noise floor alone} (no
    multi-access interference), of the 12 authentic signals, there are two
    signals with $C/N_0 \approx 55$ dB-Hz (PRNs 13 and 23), two others with
    $C/N_0 \approx 53$ dB-Hz (PRNs 3 and 16), and three others above
    $C/N_0 \approx 50$ dB-Hz, with many of the remaining signals in the mid-
    to high-40 dB-Hz range.  Note that the \emph{measured} $C/N_0$ values are
    approximately 3 dB lower than this due to multi-access interference. (In a
    typical receiver, measured $C/N_0$ is actually $C/(N_0 + M_0)$.)

    Note also that, for TEXBAT Scenario 3, $\eta_i = 1.3$ for all $i$.  Thus,
    although Scenario 3 is advertised as a ``matched power'' scenario, it is
    only approximately so.  A truly matched scenario would have $\eta_i = 1$
    for all $i$.

    The reviewer is correct that some wideband quantization noise and aliasing
    due to the rudimentary RF front-end enters the 2 MHz band around L1 for
    TEXBAT Scenario 3.  But the power of such noise is small compared to the
    power in the spoofing signals of Scenario 3.  The quantization and
    aliasing noise does tend to make TEXBAT Scenarios 1-6 easier to detect
    than a more sophisticated spoofing attack would be (because they have
    slightly higher power).  But TEXBAT Scenario 7 does not suffer from such
    quantization and aliasing noise, nor do any of the Monte-Carlo-simulated
    scenarios against which the PD detector was tested.

    In any case, to avoid any confusion, and to avoid distracting the reader
    with excuses for the rudimentary front end used in TEXBAT Scenarios 1
    through 6, we have changed Figs. 3 and 4 to reflect the spoofing scenario
    from TEXBAT Scenario 7, which is free of aliasing or quantization noise.
    We have also re-drawn the nominal (clean) plots to make them clearer.

    Furthermore, to call attention to the extra noise in TEXBAT Scenarios 2
    through 6, we have modified the second paragraph of VII.A to read

    \begin{quotation}
      Note that {\tt tb2}--{\tt tb6} exhibit a modest amount of quantization
      and aliasing noise that makes them easier to detect than would be
      expected based on this paper's models, whereas {\tt tb7} is free from
      such extraneous noise.  {\tt tb7} is also special in that the spoofer
      exercises control of $\Delta \theta$, permitting nulling, as described
      in Section V-A, in the early stages of the attack.  In all other TEXBAT
      recordings, the spoofer controlled $\eta$ and $\Delta \tau$ but left
      $\Delta \theta$ at an arbitrary constant value ({\tt tb3}, {\tt tb4},
      and {\tt tb6}) or allowed it to ramp consistent with the pull-off rate
      ({\tt tb2} and {\tt tb5}).
    \end{quotation}

{\textit{Another area that could be made clearer is that the aim of
      this paper is not
so much to detect spoofing per se, but rather to
      detect a carry-off attack.
Thus H2 is the hypothesis that a carry-off
      attack is underway. This pre?supposes
that the receiver is initially
      tracking true GPS signals. This is a small point
but it is important to
      be clear on this.}}

 We agree that this clarification is warranted, and have modified the
  abstract, the introduction, and the conclusion to make clear that our
  spoofing detection applies to carry-off-type attacks.

\subsection*{Minor Comments:}

{\textit{Introduction: "... no civil GNSS signals yet incorporate cryptographic modulation...
" \\
      perhaps better to say "... no open civil GNSS signals ..." as the
      Galileo
Commercial Service is indeed a civil GNSS signal incorporating
      cryptographic
modulation.}}

  {We have incorporated this clarification in our revised manuscript.}

{\textit{P. 11: Under H2: Spoofing. Here you do not permit advance of the
      spoofed signal
with respect to the authentic signal, why?  Also, you
      state that spoofing where
the time difference exceeds two chips "is
      more properly classified as jamming".
I would argue that under certain
      circumstances this could well be classified
as very successful
      spoofing.}}

   Thank you for pointing this out.  We unwittingly left off the absolute
    value bars around $\Delta \tau$ here and in the description of $H_3$.  We
    have now fixed this, indicating that spoofing can be either advanced or
    delayed wrt the authentic signal.  We have also modified the text to
    justify our classification of spoofing as jamming when $|\Delta \tau|$ is
    sufficiently large:

    \begin{quotation}
      These bounds make sense because the spoofing signal must be at least as
      powerful as the authentic signal for reliable spoofing, and because
      spoofing whose $|\Delta \tau|$ value exceeds $\Delta \tau_1 = 2\tau_c$
      is no longer classified as carry-off-type spoofing, since the
      interference is uncorrelated with the authentic signal.  Instead, for
      $|\Delta \tau| \geq \Delta \tau_1$, the attack is classified as jamming.
    \end{quotation}

    We have left the assumed distribution of $\Delta \tau$ under $H_2$ as it
    was, since the worst case attack occurs when the spoofer mimics multipath.

{\textit{Figure 10 ? there's a floating exclamation mark (!) after
      the figure ? most
likely a TeX float placement command gone
      wrong...}}

  {{We have fixed this.  You were correct that it was a placement command
        gone wrong.}}

\section*{Reviewer 2}
\subsection*{Suggestions for Improvement:}

{\textit{The paper is generally well written and organized; however, the
      notation used is sometimes distracting. The proposed power-distortion
      detector is very interesting in that it can detect both high-power and
      low-power spoofing attempts. The authors acknowledge that the monitor is
      ineffective in cases where the spoofer is able to block or null the
      authentic GPS signals (which is difficult to accomplish). However, in
      the results section, while discussing Table 3, it is stated that the
      proposed detector is most effective during the initial spoofer
      pull-off. Beyond a certain point, the distortion will subside, and the
      monitor's effectiveness would decrease.}}

   Agreed.  We have modified the paper's abstract, introduction, and
    conclusion to emphasize the fact that its spoofer detection is designed to
    work for carry-off spoofing attacks; in particular, during the initial
    stages of a carry-off attack when distortion remains significant.

{\textit{It would be interesting to see how very slow spoofer pull-off
      would affect the monitor, and what the resulting minimum detectable
      error would be. It appears unlikely that the proposed monitor can be
      completely fooled by a slow pull-off rate. However, the experimental
      results section does not mention details regarding the spoofer's
      pull-off rate. If it is very slow, is seems possible that the detector
      would initially classify the spoofing as multipath. Then, beyond a
      certain minimum detectable error, the detector would correctly identify
      the spoofing attempt. It is possible that the reason why results in
      Table 3 show no instances where spoofing is classified as multipath is
      because of a sufficiently quick pull-off rate.}}

   The reviewer is correct that close-in (small $|\Delta t|$),
    low-powered spoofing could fool the PD detector into declaring multipath
    ($H_1$).  However, a slow pull-off would afford the PD detector
    \emph{more} opportunities to observe distortion and declare the attack as
    spoofing than a fast pull-off attack.  We believe that the primary reason
    no empirical spoofing scenarios were mis-classified as multipath is that
    they were all designed to be \emph{effective} spoofing attacks, with
    sufficiently large $\eta$ to effect pull-off of the tracking points.
    Thus, while they were sometimes confused for jamming, their received power
    was too strong to be confused with multipath.

    The three references to TEXBAT data given in Section VII-A offer details
    on the pull-off rates of each attack.  In particular, {\tt tb7} has a
    particularly slow pull-off, whereas {\tt tb2} throuth {\tt tb6} are faster.

\subsection*{Minor Comments:}

{\textit{Variable `C' is used to represent the C/A code, and later used
      to represent the cost function}}

  {{We have modified $C(t)$ to read $C_r(t)$ throughout to avoid this confusion.}}

{\textit{letter 'j' is used to denote the imaginary unit on pg. 5,
      as well as a time index (pg. 7), and parameter space index
      (pg. 22) }}

  {{ Thanks for pointing this out. However, in this context, we do feel that
      the meaning of our notation is clear. That is, $j$ is often used as an
      indexing variable, and we have already used $i$ and $k$ elsewhere in the
      paper. Likewise, either $i$ or $j$ are also traditionally used to
      indicate imaginary components. Given that there is no overlap within the
      same expression, we'd prefer to keep the notation as is in this case.}}

{\textit{ Two different notations for conditional probability are used
      throughout the paper, both the conventional notion, P(A|B), and
      subscript notion, $P_B(A)$. At times, the subscript notion can become
      distracting, because it is used to denote conditional as well as
      marginal distributions (pg. 9 for example). For example, on page 16, the
      notation used for the conditional density of $D_k$, defined as
      $p_{Dk}(x | \theta)$, is confusing because it appears that both
      notations are used simultaneously. Adhering to the conventional notion,
      $P(A|B)$, for conditional probabilities would improve readability, and
      eliminate extraneous definitions (on pg. 8 for example). }}

  {{ Thank you for this suggestion.  We agree and have
        updated our notation in the revision.  $p_\theta(z_k)$ is now written
        as $p(z_k | \theta)$, and $p_i(z_k)$ as $p(z_k | \Theta \in
        \Lambda_k)$.  We have retained the notation $w_i(\theta)$ for
        compactness and because there is no risk of ambiguity. }}

{\textit{When the cost function is introduced on page 8, the
      definition C[i,$\theta$] is puzzling because it is simply defined
      as the cost of choosing hypothesis i, given $\theta$ belonging to
      $\Lambda$ (the entire parameter space). This initially made it
      seem that the cost is uniform across the entire parameter
      space. Later in the paper, section 6 helps clarify the definition
      by showing two different types of cost function. Uniform costs are
      discussed, as well as non-uniform costs, for various cases of
      $\theta$ in $\Lambda_i$. Additional explanation of the cost
      function definition on page 8 would be beneficial. }}

    Thank you for this pointing out this need.  We agree and have added
    the following language around the introduction of the cost function:

        \begin{quotation}
          Let $C[i,{\theta}]$ be the cost of choosing $H_i$ when
          ${\theta} \in \Lambda$ is the actual parameter vector.  Note that
          this function is sensitive to a particular value of $\theta$, which
          makes it more general than one that simply assigns a unform cost for
          choosing $H_i$ when $\theta \in \Lambda_j$.  A later section will
          introduce various embodiments of $C[i,{\theta}]$.
        \end{quotation}

\section*{Reviewer 3}
\subsection*{Suggestions for Improvement:}

{\textit{Figure 2 can be improved by using some 3D effect on the
      correlation functions.}}

   Fig. 2 already includes some features that help the reader view it in
    3 dimensions.  First, it offers a perspective view of a 3D coordinate
    frame in $I$, $Q$, and $\tau$.  Second, in the upper-right inset it allows
    the viewer to appreciate the angular difference between $\xi_{Ak}$ and
    $\xi_{Ik}$ by offering a view along the $\tau$ axis.

{\textit{Please explain why the signal models in section II did not
      include a frequency component. With interference or any strong secondary
      signal, it is possible for the tracking loop to diverge which results in
      tracking frequency error. If the error is sufficiently small, it may
      result in a time varying phase.}}

   The models in Section II do not include a carrier frequency component,
    since they are expressed in a complex baseband representation.  But they
    do include a Doppler frequency component: $\theta_A$ and $\theta_I$ are
    time varying, as noted in the original text:
    \begin{quotation}
      $P_{\rm A}$, $\tau_{\rm A}$, and $\theta_{\rm A}$ are assumed to vary
      with time, but their time dependency is suppressed for notational
      simplicity.
    \end{quotation}
    Also, the relative time variation between $\theta_{\rm A}$ and
    $\hat{\theta}$ -- the variation that could cause tracking stress -- is
    explicitly noted later on by the notation $\theta_{{\rm A}k}$.

{\textit{p6: Technically, R(tau) is a cross correlation function.}}

  {{Agreed.  But the technical distiction was already noted in the
        original manuscript:}}
    \begin{quotation}
      The function $R(\tau) = E[C_r(t)C_{\ell}(t - \tau)]$, often called the
      autocorrelation function of $C_r(t)$ even though, strictly speaking,
      $C_{\ell}(t)$ may be slightly different from $C_r(t)$, approximates the
      interaction between $C_r(t)$ and $C_{\ell}(t)$ over the correlation and
      accumulation operations.
    \end{quotation}

{\textit{section VIII: How is the performance compared against
      other published methods?}}

   Of the various signal authentication techniques mentioned in the
    paper's introduction (which offers an up-to-date summary of techniques),
    this paper's approach can be fairly compared only with a small set of
    practical near-term GNSS signal authentication techniques that do not
    require changes to GNSS signals-in-space, are receiver-autonomous,
    low-cost, require no additional hardware, and can be implemented via a
    software or firmware update. Existing techniques in this category include
    monitoring the total received power via the Automatic Gain Control (AGC)
    setpoint~\cite{akos2012s}, and monitoring autocorrelation profile
    distortion~\cite{ledvina2010cgr, cavaleri2010dscc, huang2016gnss}.  These
    tecniques can be viewed as special cases of the current paper's technique.
    The ``blindspot'' from which these techniques suffer due to their ignoring
    one or the other of received power or correlation function distortion can
    be appreciated emph{a priori}: A received power monitor that ignores
    correlation distortion may not detect a low-power spoofer. Moreover,
    because a power-monitoring-only technique does not distinguish between
    spoofing and jamming, its alarm rate can become intolerable in urban areas
    where so-called personal privacy devices (PPDs, small GNSS
    jammers)~\cite{mitch2012kyejam} are common.  For their part, the
    distortion-monitoring approaches in \cite{ledvina2010cgr,
      cavaleri2010dscc, huang2016gnss} ignore total received power, and thus
    can be fooled by a spoofer transmitting with a significant power advantage
    over the authentic signals, which, by action of the AGC, forces the
    authentic signals under the noise floor, leaving a distortion-free
    correlation function~\cite{humphreys2011ds}.

    This \emph{a priori} comparison is included in the paper's original
    manuscript.  It is also clear from the optimal decision regions shown in
    Fig. 7 that ignoring one or the other of $P_k$ or $D_k$ (as the prior work
    in this area does), would be sub-optimal.  Therefore, we believe that the
    paper convincingly establishes that its approach is superior to other
    approaches in the relevant category for comparison.

{\textit{More details on the test data are expected.}}

  {{Thank you. An additional reference for the TEXBAT data set was
        included.  Between the three references to the TEXBAT data, and the
        two references on the ``RNL Multipath and Interference Recordings,''
        we believe the test data are now adequately detailed.}}

{\textit{I am not sure that the paper includes enough results to
      show how well the proposed method functions with weak signals. For
      example, if the target signal is weak, interference from other
      GNSS signals may also confuse the detector, since it can be
      observed as multipath or interference. It may trigger more false
      alarms.}}

  {{The reviewer raises a good point here: it is possible that near-far
        effects (interaction of weak signals with very strong signals due to
        imperfect orthogonality of their spreading codes) could cause
        distortion that would lead to false spoofing alarms.  The reviewer
        encourages testing to determine if the proposed method works in
        near-far cases with a large gap between the power of the strongest and
        weakest signal.  In fact, the clean data set {\tt cd0}, {\tt wd0}, and
        {\tt wd1} do include a fairly wide range of signal powers.  The most
        powerful signals show $C/N_0 \approx 52$ dB-Hz, while the weakest
        signals are at 35 dB-Hz.  This constitutes 17-dB difference in signal
        power, yet no false spoofing alarms were raised (cf. Table III).  An
        even wider dynamic range could be tested, but the reviewer is
        requested to keep in mind that the difficulty of distinguishing
        multipath from low-power spoofing when severe shadowing is present was
        already treated in Section III:

        \begin{quotation}
          On the other hand, the severe shadowing experienced by mobile
          receivers can cause $\eta > 1$, especially in urban environments
          where surrounding buildings simultaneously attenuate and reflect
          authentic signals \cite{steingass2004measuring}.  In these
          situations, there is no practical upper limit on $\eta$ because the
          direct-path authentic signal may be attenuated by 50 dB or more. It
          is not possible to reliably distinguish multipath from low-power
          spoofing in such circumstances.

          One can avoid this difficulty by applying this paper's detection
          test only when the authentic signal $r_{\rm A}(t)$ is received
          without severe shadowing, in which case the probability that
          $\eta > 1$ under $H_1$ again becomes negligible. In particular, the
          multipath model developed below assumes that $r_{\rm A}(t)$ is
          attenuated less than 6 dB by shadowing.  In practice, one simply
          excludes cases in which the received power $P_k$ is not unusually
          high yet the measured carrier-to-noise ratio $C/N_0$ drops by more
          than 6 dB from its modeled value for an unobstructed authentic
          signal.
        \end{quotation}

        In other words, our approach avoids the effects of severe shadowing by
        applying the PD detector only to signals that have been attenuated
        less than 6 dB by shadowing.  With this exclusion, and with the
        false-alarm-free performance of the paper's methods on the three $H_0$
        data sets, in additions to thorough simulation testing in Section VI,
        we believe that the spoofing false alarm rate under $H_0$ will be
        negligible.}}

{\textit{For the power metric introduced in IV.A, it is not
      entirely clear how it can be implemented. How precise does
      the power measurement has to be in order for this method
      to be applicable? In practice, we use carrier to noise
      ratio to gauge signal strength. Depending on the
      implementation choice, it may not be very precise for weak
      signals, or signals with interference. }}

   Details on how to measure $P_k$ in practice are included in this
    paragraph from Section IV.A:

    \begin{quotation}
      Receivers with sufficient dynamic range in the discrete samples produced
      from $r_{\rm C}(t)$ do not require an AGC and so can compute received
      power directly by averaging the squared modulus of these samples.  Let
      $W_{P}$ be the bandwidth over which $P_k$ is to be measured.  If $W_P$
      is narrower than the receiver front-end's native bandwidth, then
      $r_{\rm C}(t)$ must be filtered to isolate the desired spectral
      interval.  Fig. 3 shows 2- and 8-MHz bands for $P_k$,
      with Fig. 4 showing the corresponding power time
      histories.  In a receiver whose front end is equipped with an AGC, $P_k$
      is measured indirectly through the AGC setpoint \cite{akos2012s}.  In
      this case $W_{P}$ is equivalent to the front-end noise-equivalent
      bandwith.
    \end{quotation}

    Details on measuring the power indirectly through the AGC setpoint are
    found in \cite{akos2012s}, referenced within this paragraph.

      Note that measuring aggregate received power is different from measuring
      $C/N_0$.  Consider that, if effective $N_0$ rises in a spoofing attack
      commensurate with carrier power $C$, then $C/N_0$ will remain constant
      even as $P_k$ rises.

      How accurately does $P_k$ need to be measured to be useful?  We have
      modified the manuscript in two places to indicate that the paper's
      detection test assumes $\sigma_P = 0.4$ dB.  This is roughly four times
      the standard deviation due to natural variation in a high quality
      receiver in a quiet RF environment (see the top panel in Fig. 4 and the
      discussion within the text).  Receivers whose measurement of $P_k$ is
      significantly less precise than this, due to quantization or inaccuracy
      in the AGC set point-to-$P_k$ mapping, will see their ability to detect
      low-power spoofing and jamming reduced.

      Note that \cite{akos2012s} showed that the precision of the AGC
      measurement as a proxy for in-band power is much better than the natural
      variation in the AGC (see Fig. 7 in \cite{akos2012s}).  This indicates
      that $P_k$ can be reliably obtained from the AGC measurement (to within
      a constant offset that can be calibrated in a quiet RF environment).

{\textit{Eq 10 assumes that the code phase can be observed
      by maximizing correlation function. Ideally this
      assumption would be nice to have. However, with weak
      signals or limited bandwidth, it is usually not practical
      to do that. The early-late code discriminator is likely to
      be more sensitive than the peak point in some
      scenarios. What is the sensitivity level of the detector
      against errors made in this step? }}

   This paper's assumed recipe for determination of $\hat{\tau}$ is given
    by \[ \hat{\tau} = \arg \max_\tau \xi_k(\tau)  \]  But this recipe is
    only used for modeling purposes.  As the reviewer notes, in practice the
    recipe will be some type of early-late code discriminator, possibly with
    multiple early and late taps.

    The actual receiver used to generate the observables that were processed
    by the PD detector to yield the results in Section VII, the GRID
    science-grade software-defined receiver \cite{lightsey2013demonstration},
    does in fact use an early-minus-late discriminator to obtain $\hat{\tau}$.
    The fact that the PD detector worked well when using the GRID-produced
    observables---even in the case of signals as weak as $35$
    dB-Hz---indicates that the method is not highly sensitive to how
    $\hat{\tau}$ is obtained.

\bibliographystyle{IEEEtran} 
\bibliography{pangea}
\end{document}

### 2022-10-29 paper.tex
init

\documentclass[10pt,journal,twocolumn]{IEEEtran} % add 'draft' to suppress figures
% Double-spaced, single-column format for IEEE journal submission
% \documentclass[12pt,journal,draftclsnofoot,onecolumn]{IEEEtran} 
\usepackage{wrapfig,booktabs,fancyhdr,amsmath,amsfonts,tabularx,numprint}
\usepackage{cite,bm,bbm,amssymb,amsthm,url,multirow,times,enumitem,comment}
\usepackage{mathtools,siunitx,balance,tikz,adjustbox,graphicx,array}
\usepackage[font=footnotesize]{subcaption}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage[colorlinks=true, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
\newcommand{\vb}{\boldsymbol}
\newcommand{\vbh}[1]{\hat{\boldsymbol{#1}}}
\newcommand{\vbc}[1]{\check{\boldsymbol{#1}}}
\newcommand{\vbb}[1]{\bar{\boldsymbol{#1}}}
\newcommand{\vbt}[1]{\tilde{\boldsymbol{#1}}}
\newcommand{\vbs}[1]{{\boldsymbol{#1}}^*}
\newcommand{\vbd}[1]{\dot{{\boldsymbol{#1}}}}
\newcommand{\abs}[1]{\left|{#1}\right|}
\newcommand{\by}{\times}
\newcommand{\tr}{\mathsf{T}}
\newcommand{\sfrac}[2]{\textstyle \frac{#1}{#2}}
\newcommand{\ba}{\begin{array}}
\newcommand{\ea}{\end{array}}
\newcommand{\sinc}{\text{sinc}}
\newcommand{\define}{\triangleq}
%\newcommand{\define}{\coloneqq}
\newcommand{\cnr}{C/N_0}
\newcommand{\sgn}{\text{sgn}}
\renewcommand{\Re}{\mathbb{R}}
\renewcommand{\Im}{\mathbb{I}}
\newcommand{\E}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\brm}[1]{\bm{\text{#1}}}
\newcommand{\round}[1]{\ensuremath{\lfloor#1\rceil}}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\DeclareMathOperator*{\argmin}{\arg\!\min}
\DeclareMathOperator*{\argmax}{\arg\!\max}

% Macros for this paper
\newcommand{\Fc}[1]{\ensuremath{F_{\text{c}#1}}}
\newcommand{\Fs}{\ensuremath{F_\text{s}}}
\newcommand{\Fsr}{\ensuremath{F_\text{sr}}}

% \topmargin = 0 mm 
% \oddsidemargin = -1 mm 
% \evensidemargin = -1 mm
% \headheight = 0 mm 
% \headsep = 8 mm 
% \textheight = 220 mm 
% \textwidth =170 mm 
% \parindent = 0 mm
% \parskip = 4 mm

\begin{document}
%%% ----------------------------------------------------------Front Matter
\title{An Exegesis of Ontological Hermeneutics}

\author{
  \IEEEauthorblockN{Gandalf the White\IEEEauthorrefmark{1}, Radagast the
    Brown\IEEEauthorrefmark{2}, Todd E. Humphreys\IEEEauthorrefmark{1}} \\
  \IEEEauthorblockA{\IEEEauthorrefmark{1}\textit{Department of Aerospace
      Engineering and Engineering Mechanics, The University of Texas at Austin}} \\
  \IEEEauthorblockA{\IEEEauthorrefmark{2}\textit{Department of Elven Lore, The
    University of East Rivendell}}
}

\maketitle

\begin{abstract}
  The abstract is a one-paragraph summary of the whole paper.  Out of respect
  for busy readers, it gets right to the point.  The first 2-3 sentences
  unpack the title and motivate the paper.  The remainder briefly offers
  context (if necessary) and further describes the paper's method, explaining
  how it fills a gap in the existing literature.  The abstract reveals all the
  paper's main results in language that is bold but truthful.

An example for a paper titled ``Deep urban unaided precise GNSS vehicle
positioning'': \\ \\

This paper presents the most thorough study to date of vehicular carrier-phase
differential GNSS (CDGNSS) positioning performance in a deep urban setting
unaided by complementary sensors.  Using data captured during approximately 2
hours of driving in and around the dense urban center of Austin, TX, a CDGNSS
system is demonstrated to achieve 17-cm-accurate 3D urban positioning (95\%
probability) with solution availability greater than 87\%.  The results are
achieved without any aiding by inertial, electro-optical, or odometry sensors.
Development and evaluation of the unaided GNSS-based precise positioning
system is a key milestone toward the overall goal of combining precise GNSS,
vision, radar, and inertial sensing for all-weather high-integrity
high-absolute-accuracy positioning for automated and connected vehicles.  The
system described and evaluated herein is composed of a densely-spaced
reference network, a software-defined GNSS receiver, and a real-time kinematic
(RTK) positioning engine.  A performance sensitivity analysis reveals that
navigation data wipeoff for fully-modulated GNSS signals and a dense reference
network are key to high-performance urban RTK positioning.  A comparison with
existing unaided systems for urban GNSS processing indicates that the proposed
system has significantly greater availability or accuracy.
\end{abstract}

\begin{IEEEkeywords} 
urban vehicular positioning; CDGNSS; low-cost RTK positioning.
\end{IEEEkeywords}

%%% -------------------------- Select preprint or "for submission" version
\newif\ifpreprint
\preprintfalse
%\preprinttrue

\ifpreprint

\pagestyle{plain}
\thispagestyle{fancy}  
\fancyhf{} 
\renewcommand{\headrulewidth}{0pt}
\rfoot{\footnotesize \bf October 2020 preprint of paper submitted for review} \lfoot{\footnotesize \bf
  Copyright \copyright~2020 by Lakhsay Narula \\ and Todd E. Humphreys}

\else

\thispagestyle{empty}
\pagestyle{empty}

\fi

%%% ----------------------------------------------------------------------

\section{Introduction}
In the introduction:
\begin{itemize}
\item {\bf Background:} Set the stage in a compelling way to draw the reader
  in.
\item {\bf Opportunity/Need:} Identify a need or an opportunity.
\item {\bf The status quo:} Review the state of the art.  
\item {\bf Gap:} Mind the gap!  Convince your reader that there is an
  important deficiency (gap) in the existing literature or practice, and that
  your paper will save the day by filling in/bridging this gap.  Otherwise,
  why should your reader take time to read the paper?
\item {\bf Vision:} Invite the reader to imagine the way the world could be if
  only the gap could be bridged.
\item {\bf Contributions:} Clearly enumerate the paper's contributions.
\end{itemize}

What follows is an example introduction taken from
\cite{humphreys2019deepUrbanIts}. \\ \\

\IEEEPARstart{F}{uture} Vehicle-to-vehicle (V2V) and vehicle-to-infrastructure
(V2I) connectivity will permit vehicles to relay their positions and
velocities to each other with millisecond latency, enabling tight coordinated
platooning and efficient intersection management.  More ambitiously, broadband
V2V and V2I enabled by 5G wireless networks will permit vehicles to share
unprocessed or lightly-processed sensor data. \emph{Ad hoc} networks of
vehicles and infrastructure will then function as a single sensing organism.
The risk of collisions, especially with pedestrians and cyclists---notoriously
unpredictable and much harder to sense reliably than vehicles---will be
significantly reduced as vehicles and infrastructure contribute sensor data
from multiple vantage points to build a blind-spot-free model of their
surroundings.

Such collaborative sensing and traffic coordination requires vehicles to know
and share their own position.  How accurately?  The proposed Dedicated Short
Range Communications (DSRC) basic safety message, a first step in V2V
coordination, does not yet define a position accuracy requirement, effectively
accepting whatever accuracy a standard GNSS receiver provides
\cite{kenney2011dedicated}.  But automated intersection management
\cite{fajardo2011automated}, tight-formation platooning, and unified
processing of sensor data---all involving vehicles of different makes that may
not share a common map---will be greatly facilitated by globally-referenced
positioning with sub-30-cm accuracy.

Poor weather also motivates high-accuracy absolute positioning.  Every
automated vehicle initiative of which the present authors are aware depends
crucially on lidar or cameras for fine-grained positioning within their local
environment.  But these sensing modalities perform poorly in low-visibility
conditions such as a snowy whiteout, dense fog, or heavy rain.  Moreover,
high-definition 3D maps created with lidar and camera data, maps that have
proven crucial to recent progress in reliable vehicle automation, can be
rendered dangerously obsolete by a single snowstorm, leaving vehicles who rely
on such maps for positioning no option but to fall back on GNSS and radar to
navigate a snow-covered roadway in low-visibility conditions.  When, as is
often the case on rural roads, such snowy surroundings offer few
radar-reflective landmarks, radar too becomes useless.  GNSS receivers operate
well in all weather conditions, but only a highly accurate GNSS solution,
e.g., one whose absolute errors remain under 30 cm 95\% of the time, could
prevent a vehicle's drifting onto a snow-covered road's soft shoulder.  Code-
and Doppler-based GNSS solutions can be asymptotically accurate (averaged over
many sessions) to better than 50 cm, which may be adequate for digital mapping
\cite{narula2018accurate}, but they will find it challenging to meet a 30 cm
95\% stand-alone requirement, even with modernized GNSS offering wideband
signals at multiple frequencies.

Carrier-phase-based GNSS positioning---also referred to as precise GNSS
positioning even though it actually offers absolute accuracy, not just
precision (repeatability)---can meet the most demanding accuracy requirements
envisioned for automated and connected vehicles, but has historically been
either too expensive or too fragile, except in open areas with a clear view of
the overhead satellites, for widespread adoption.  Coupling a carrier-phase
differential GNSS (CDGNSS) receiver with a tactical grade inertial sensor, as
in \cite{petovello2004benefits,scherzinger2006precise,
  zhangComparisonWithTactical2006,kennedy2006architecture} enables robust
high-accuracy positioning even during the extended signal outages common in
dense urban areas.  But GNSS-inertial systems with tactical-grade inertial
measurement units (IMUs) cost tens of thousands of dollars and have proven
stubbornly resistant to commoditization.  Coupling a GNSS receiver with
automotive- or industrial-grade IMUs is much more economical, and
significantly improves performance, as shown in \cite{li2018high}.  But such
coupling only allows approximately 5 seconds of complete GNSS signal blockage
before the IMU no longer offers a useful constraint for so-called integer
ambiguity resolution \cite{evaluationLowCostMems2006Godha}, which underpins
the fastest, most accurate, and most robust CDGNSS techniques, namely,
single-baseline RTK, network RTK, and PPP-RTK
\cite{teunissen2015review,cui2017rtkforCAV}.

Previous research has suggested an inexpensive technique for robustifying RTK
positioning: tightly coupling carrier-phase-based GNSS positioning with
inertial sensing and vision \cite{shepard2014fusion,pesyna2015dissertation}.
Such coupling takes advantage of the remarkable progress in high-resolution,
low-cost cameras within the intensely competitive smartphone market.  The
current authors are engaged in developing a high-integrity RTK-vision system
for high-accuracy vehicular positioning in rural and urban environments.
Further coupling with radar will make the system robust to low-visibility
conditions.

As a step toward this goal, it is of interest to evaluate the performance of
stand-alone RTK techniques---those unaided by IMUs, odometry, or vision---in
urban environments.  Such a study will reveal why and when aiding is
necessary, and how an RTK positioning system might behave if aiding were
somehow impaired or unavailable, whether due to sensor faults or, in the case
of exclusive visual aiding, poor visibility conditions.

Little prior work has explored unaided vehicular RTK performance in urban
environments, no doubt because performance results have historically been
dismal. Short-baseline RTK experiments between two vehicles in
\cite{ong2009assessment} revealed that multi-frequency (L1-L2) GPS and Glonass
RTK yielded poor results in residential and urban environments.  Only along a
mountain highway with a relatively clear view of the sky was availability
greater than 90\% and accuracy better than 30 cm.  RTK positioning in downtown
Calgary was disastrous, with less than 60\% solution availability and RMS
errors exceeding 9 meters.

More recently, Li et al. \cite{li2018high} have shown that, with the benefit
of greater signal availability, unaided professional-grade dual-frequency GPS
+ BDS + GLONASS RTK can achieve correct integer fixing rates of 76.7\% on a
1-hour drive along an urban route in Wuhan, China.  But Li et al. do not
provide data on the incorrect fixing rate, nor a full error distribution, so
the significance of their results is difficult to assess.

Recent urban RTK testing by Jackson et al. \cite{jackson2018assessmentRtk}
indicates that no low-to-mid-range consumer RTK solution offers greater than
35\% fixed (integer-resolved) solution availability in urban areas, despite a
dense reference network and dual-frequency capability.  A key failing of
existing receivers appears to be their slow recovery after passing under
bridges or overpasses.

This paper describes and evaluates an unaided RTK positioning system that has
been designed for vehicular operation in both rural and urban environments.
Preliminary performance results were published in a conference version of this
paper \cite{humphreys2018urbanStandAloneRTKplans2018}.  The current paper
improves on the conference version in four ways: (1) the test route is both
more challenging and more comprehensive, (2) a proper independent ground truth
trajectory is used as the basis of error evaluation, (3) data modulation
wipeoff for improved carrier tracking robustness is applied not only on GPS L1
C/A signals, as previously, but now also on SBAS L1 signals, and (4) the
performance benefit of vehicle GNSS antenna calibration is assessed.

This paper's primary contributions are (i) a demonstration of the performance
that can be achieved with a low-cost software-defined unaided RTK GNSS
platform in a dense urban environment, and (ii) an evaluation of the relative
importance of various factors (e.g., data bit wipeoff, age of reference data,
rover antenna calibration, reference network density) to the overall system
performance.

To stimulate further innovation in urban precise positioning, all data from
this paper's urban driving campaign have been posted at
\url{http://radionavlab.ae.utexas.edu} under ``Public Datasets,'' including
wideband (10 MHz) intermediate frequency samples from both the reference and
rover antennas, RINEX-formatted rover and reference observables, and the
ground truth trajectory.

\section{Tables}
Captions should appear above tables.  Use the full word Table to refer to a
table in your paper; e.g., ``Results of the study are shown in Table I.''
Follow the format of Table \ref{tab:Chi2Fits} and Table
\ref{tab:ValidationTable2}.

\begin{table}[ht]
  \centering
  \caption{Chi-square Values for Fits to Nakagami-m and Rice Distributions}
  \begin{tabular}[c]{cccc}
    \toprule
    Data Source &  Sets, DOF & Nakagami-m & Rice \\ \midrule
    Wideband UHF & 79, ~~8 & 11.8 $\pm$ 8.8 & 9.0 $\pm$ 4.3 \\
    GPS L$_1$ & 33, ~~7 & 8.42 $\pm$ 5.9 & 7.7 $\pm$ 5.7 \\ \bottomrule
  \end{tabular}
  \label{tab:Chi2Fits}  
\end{table}

\begin{table}[ht]
  \centering
  \caption{Scintillation Effects Comparison:~~Empirical Truth Data}
  \begin{tabular}[c]{ccccccc}
    \toprule
    \multicolumn{3}{c}{Parameters}  
    & \multicolumn{2}{c}{Truth Scint.} 
    & \multicolumn{2}{c}{Synthetic Scint.} \\
    \cmidrule(r){1-3} \cmidrule(r){4-5} \cmidrule(r){6-7}
    $S_4$& $\tau_0$ (s) & $T$ (s) & $\sigma_{\varphi}$ (deg) 
    & $N_s$ & $\sigma_{\varphi}$ (deg) & $N_s$ \\  \midrule
    0.87 & 0.18 & 200 & 16.4 & 32 & 17.5 $\pm$ 0.5 & 35.9 $\pm$ 4.7 \\ 
    1.0 & 0.36 & 265 & 14.1 & 37 & 15.0 $\pm$ 0.5 & 41.6 $\pm$ 5.9 \\  
    0.69 & 0.18 & 174 & 12.7 & 12 & 11.8 $\pm$ 0.9 & 5.6 $\pm$ 1.6 \\ 
    0.87 & 0.26 & 225 & 11.6 & 23 & 12.7 $\pm$ 0.5 & 19.2 $\pm$ 4.6 \\
    0.61 & 0.47 & 162 & 3.96 & 0 & 3.63 $\pm$ 0.2 & 0.10 $\pm$ 0.3 \\
    0.96 & 0.09 & 81 & 28.5 & 60 & 32.7 $\pm$ 1.0 & 69.4 $\pm$ 5.8 \\
    0.95 & 0.26 & 123 & 14.1 & 21 & 15.6 $\pm$ 0.5 & 19.8 $\pm$ 3.7 \\
    0.51 & 0.71 & 138 & 2.12 & 0 & 1.60 $\pm$ 0.1 & 0 $\pm$ 0~~~ \\
    \bottomrule
  \end{tabular}
  \label{tab:ValidationTable2}  
\end{table}

\section{Figures}
\begin{enumerate}
\item Use Fig. instead of Figure to refer to a figure (per IEEE style guide).
  E.g., ``Fig. (4) illustrates this point by plotting $x$ vs. $y$.''
\item Only use color in figures if necessary.  Many figures look better in
  print if done in grayscale.  See the example figures in the writing templates.
\item Try to make figure captions sufficient, or nearly so, to describe the
  figure.  For example, the caption should explain what the various lines or
  enclosed regions in a figure mean.  Push description from the text to the
  caption to avoid duplication.
\item Avoid legends in figures; prefer labels that point with lines directly
  to the object being labeled.
\item Make figures as small as they can be while remaining clearly legible and
  conveying their message.  Often, you can use {\tt subplot(211)} in Matlab to
  cut a figure height in half.
\item Figures should not have titles on top.  All relevant information should
  be put in the caption.
\item Ensure that each figure's x and y axis labels, and all internal labels,
  are in the same or similar font as the text and are large enough to read
  easily.  A rule of thumb: make labels no smaller than the caption font.  You
  can use {\tt figset.m} to change native Matlab font to one that more closely
  resembles LaTeX default font and increase its size.
\end{enumerate}

Have a look at Fig. \ref{fig:rocplot1} as an example.  Todd uses an
application called xfig to add labels and LaTeX symbols directly to eps
figures.  It's old school, but he hasn't found a better alternative for
perfectly matching the LaTeX fonts in the text.

\begin{figure}[t]
\centering
\includegraphics[width = 8.5cm]
{figs/rocplot1}
\caption{ROCs for the ML, MAP, and MMSE security code estimation strategies
  under the following scenario: $(\cnr)_s = 54$ dB-Hz,
  $(\cnr)_r = 48$ dB-Hz, $\eta = 1$, $d = 0$, $N = 400$. Top panel: High-rate
  security code with $T_w = 2~ \mu$s.  Bottom panel: Low-rate security code
  with $T_w = 20$ ms.}
\label{fig:rocplot1}
\end{figure}

Sometimes you have to take up both columns with a figure, as in
Fig. \ref{fig:rx-model}.

\begin{figure*}[t]
  \centering
  \includegraphics[width=\textwidth]{figs/correlation}
  \caption{Block diagram of the standard AGC, correlation, and accumulation
    operations in a GNSS receiver. The product of the AGC-scaled incoming
    signal $\beta(t)r(t)$ and the conjugate of the local replica
    $\ell(t,\tau)$ is accumulated over $T$ seconds to produce the discrete
    complex-valued accumulation product $\xi_k(\tau)$. For notational
    convenience, the accumulation product has been scaled by $1/T$.}
  \label{fig:rx-model}
\end{figure*}

\section{Algorithms}
Use the full word Algorithm to refer to an algorithm in your paper; e.g.,
``Algorithm~\ref{algo:correlation-search} shows the pseudocode for this
process.'' Use the following format for algorithms. Note that variables should
be written in mathematical italic font, whereas functions can be written either
in typewriter font; e.g.,
\[D_{\text m} = \texttt{toOGM} \left( \bm{p}_{\text m}^{W} - \bm{t}_k^{V},
    \delta t \right)\] or in mathematical italic font.

\begin{algorithm}
  \footnotesize
  \SetKwInOut{Input}{Input}
  \SetKwInOut{Output}{Output}
  \Input{$\bm{p}_{\text m}^W$, $\bm{p}_{{\text b, } 1:k}^B$, $\bm{t}_{1:k}^{V}$, $\phi_{1:k}^{V}$, $\sigma_t$, $\sigma_\phi$, $\delta t$, $\delta \phi$}
  \vspace{0.25em}
  \Output{$\widehat{\bm{\Theta}}$}

  \vspace{1em}

  $D_{\text m} = \texttt{toOGM} \left( \bm{p}_{\text m}^W - \bm{t}_k^{V}, \delta t \right)$

  $\tilde{D}_{\text m} = \texttt{FFT2} \left( \texttt{pad} \left( D_{\text m}, 3\sigma_t \right) \right)$

  \vspace{1em}

  $\bm{p}_{{\text b, } 1:k}^{V} = R \left( \phi_{1:k}^{V} \right) \bm{p}_{{\text b, } 1:k}^B + \bm{t}_{1:k}^{V}$

  $D_{\text b} = \texttt{toOGM} \left( \bm{p}_{{\text b, } 1:k}^{V} - \bm{t}_k^{V}, \delta t \right)$

  $\tilde{D}_{\text b} = \texttt{FFT2} \left( \texttt{pad} \left( D_{\text b}, 3\sigma_t \right) \right)$

  \vspace{1em}

  $n = 3\sigma_t / \delta t$

  $m = 3\sigma_\phi / \delta \phi$

  \For{i = -m:m}{ 
    $\Delta \phi = i \delta \phi$

    $\tilde{D}_{\text b}^{\Delta \phi} = \texttt{rotate2} \left( \tilde{D}_{\text b}, \Delta \phi \right)$

    $\mathsf{R}[i,:,:] = \texttt{IFFT2} \left( \tilde{D}_{\text m} \circ \texttt{conj} \left( \tilde{D}_{\text b}^{\Delta \phi} \right) \right) \left[ -n : n, -n : n \right]$
  }
  $\widehat{\bm{\Theta}} = \argmax \left( \mathsf{R} \right)$

  \caption{\texttt{fastGlobalAlign}}
  \label{algo:correlation-search}
\end{algorithm}

\begin{algorithm}
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\Input{$\vbh{z} \in \mathbb{R}^m$, $L \in \mathbb{R}^{m \times m}$}
\Output{$\vbc{z} \in \mathbb{Z}^m$}
$\vbh{z}_{\text c} = \vbh{z}$

\For{i = 1:m}{ 
$\check{{z}}_{i}=\lfloor\hat{{z}}_{{\text c}i} \rceil$

$\check{{\epsilon}}_{{\text c}i}=\hat{{z}}_{{\text c}i} - \check{{z}}_{i}$

\For{j = i+1:m}{
$\hat{{z}}_{{\text c}j}=\hat{{z}}_{{\text c}j} - l_{ij}\check{{\epsilon}}_{{\text c}i} $\
}
}
\caption{${\text{IB}}( \vbh{z}, L)$}
\label{algo:second-algo}
\end{algorithm}
\noindent

\section{Stylistic Conventions}
\subsection{Mathematical Notation}
\begin{enumerate}
\item Define \emph{all} mathematical notation that you introduce unless it's
  been previously defined.  Where possible, introduce notation \emph{before}
  presenting the equation involving the notation.
\item Put units in straight font, not italics or math font; e.g., cm not $cm$.
\item Write subscripts that represent indices or are just one letter long in
  the usual mathematical italic font; e.g., ${x}_{ak}$.  Write subscripts that
  are words or abbreviated words in roman type: ${x}_{\text{max}}$.
\item Defining {\tt newcommand} macros for the paper can make it easier to
  repeatedly write complicated LaTeX math expressions; e.g., $\Fc{1}$, $\Fs$,
  $\Fsr$.
\end{enumerate}

\subsection{Punctuation}
Use American Style punctuation.
\href{https://www.thepunctuationguide.com/british-versus-american-style.html}{This}
website is a good reference. Note that American style quotations include
commas and periods within the quotation, even if they are not actually part of
what was quoted; e.g.,
\begin{quotation}
  “Economic systems,” according to Professor White, “are an inevitable
  byproduct of civilization, and are, as John Doe said, ‘with us whether we
  want them or not.’”
\end{quotation}
This style is not as sensible as British English, but it's the convention on
this side of the pond.

\subsection{Abbreviations}
\begin{enumerate}
\item With few exceptions, define all abbreviations/acronyms used in
  the paper (e.g., UAV, GNSS, IMU, etc.).  Exceptions are those that have
  become so common that they're part of everyday vernacular (radar, GPS). 
\end{enumerate}

\subsection{References}
\begin{enumerate}
\item Never start a sentence with the actual numerical citation (e.g., ``[2]
  presents an optimal ...'').  Prefer instead ``Reference [2] ...'' or ``The
  authors of [2]...'' or ``Durge et al. present in [2] an optimal ...''
\end{enumerate}

\subsection{Signposts}
Busy readers rarely read a paper straight through from abstract to conclusions.
Instead, they tend to sample bits and pieces \emph{a la carte}.  For these
readers, and even for sequential readers who may need some guidance on where
they are and what is to come, it's helpful to start off each major section with
a \emph{signpost}---a brief summary of what the reader now knows and what he
can expect to learn from that section.  Sometimes the section title, perhaps in
combination with the first sub-section title, sufficies.  In other cases, a few
sentences are needed to orient the reader.  Here is an example of a
few-sentence signpost at the beginning of a section:
\begin{quotation}
  The model of road user behavior presented above applies to human
  decisions. Autonomous vehicles are likely to respond very differently,
  because they are fundamentally unable to win the game of chicken. Moreover,
  they are likely to obey the rules of the road far more than human drivers,
  particularly in safety-critical situations. This section elaborates such
  reasoning.
\end{quotation}

\section{Conclusions}
Like the abstract, this section should give a recap of the whole paper.  Write
in simple past tense.  Don't assume the reader has read the paper.

Example:\\

A real-time kinematic (RTK) positioning system tailored for urban vehicular
positioning has been described and evaluated.  To facilitate performance
comparison against similar systems, the system was tested without any benefit
of aiding by inertial or electro-optical sensors.  Over nearly 2 hours of
urban testing, including multiple passes through Austin's dense urban center,
the system achieved an 85\% probability of correct integer fix for a 2.4\%
probability of incorrect fix, resulting in 3D positioning errors smaller than
17 cm (95\%).  A performance sensitivity analysis revealed that navigation
data bit prediction on fully-modulated GNSS signals is key to high-performance
urban RTK positioning, and that a dense reference network, carrier tracking
bandwidth adaptation, and rover antenna calibration each offer a significant
integrity benefit. A comparison with existing unaided systems for urban GNSS
processing indicates that the proposed system has a significant advantage in
availability and/or accuracy.

\section*{Acknowledgments}
Example:\\

This work was supported in part by the U.S. Department of Transportation (USDOT)
under the University Transportation Center (UTC) Program Grant 69A3552047138
(CARMEN), and by affiliates of the 6G@UT center within the Wireless Networking
and Communications Group at The University of Texas at Austin.

\bibliographystyle{ieeetran} 
\bibliography{pangea}
\end{document}

### 2022-10-29 figs/figset.m

@@ -1,0 +1,14 @@
+% figset
+% Set font size of figures for printing
+
+fs = 12;
+fn = 'times new roman';
+%fn = 'helvetica';
+%fn = 'courier new';
+
+set(gca, 'fontsize', fs, 'fontname', fn);
+hy = get(gca, 'ylabel');
+hx = get(gca, 'xlabel');
+set(hy, 'fontsize', fs, 'fontname', fn);
+set(hx, 'fontsize', fs, 'fontname', fn);
+title('');

### 2023-02-07 paper.tex
PL, blocking, and shadowing

@@
-\documentclass[10pt,journal,twocolumn]{IEEEtran} % add 'draft' to suppress figures
+\documentclass[12pt]{report} % add 'draft' to suppress figures
 % Double-spaced, single-column format for IEEE journal submission
@@
 \usepackage[colorlinks=true, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
+\usepackage{geometry}
+ \geometry{
+ a4paper,
+ total={170mm,257mm},
+ left=20mm,
+ top=20mm,
+ }
+ \usepackage[acronym,nomain,nonumberlist]{glossaries}
+ \makeglossaries
+
  \newcommand{\vb}{\boldsymbol}
@@
 
+\newacronym{PL}{PL}{Path loss}
+\newacronym{LOS}{LOS}{Line of sight}
+\newacronym{NLOS}{NLOS}{Non-line of sight}
+
 % \topmargin = 0 mm 
@@
 
-\author{
-  \IEEEauthorblockN{Gandalf the White\IEEEauthorrefmark{1}, Radagast the
-    Brown\IEEEauthorrefmark{2}, Todd E. Humphreys\IEEEauthorrefmark{1}} \\
-  \IEEEauthorblockA{\IEEEauthorrefmark{1}\textit{Department of Aerospace
-      Engineering and Engineering Mechanics, The University of Texas at Austin}} \\
-  \IEEEauthorblockA{\IEEEauthorrefmark{2}\textit{Department of Elven Lore, The
-    University of East Rivendell}}
-}
+\tableofcontents
+\chapter{Path Loss Models}
 
-\maketitle
+\gls{PL} models are formulas which allow for modeling the received
+power $P_r$ given the transmit power $P_t$, distance $d$, and signal carrier
+wavelength $\lambda$. The resulting \gls{PL} is due to signal propagation
+through the environment.
 
-\begin{abstract}
-  The abstract is a one-paragraph summary of the whole paper.  Out of respect
-  for busy readers, it gets right to the point.  The first 2-3 sentences
-  unpack the title and motivate the paper.  The remainder briefly offers
-  context (if necessary) and further describes the paper's method, explaining
-  how it fills a gap in the existing literature.  The abstract reveals all the
-  paper's main results in language that is bold but truthful.
+\section{Friss Formula}
+For a transmitter and receiver spaced $d$ distance apart, the received power
+$Pr$ follows:
 
-An example for a paper titled ``Deep urban unaided precise GNSS vehicle
-positioning'': \\ \\
+\begin{align}
+  \label{eq:general_PrPt}
+  P_r = P_t \frac{1}{4\pi d} A_e
+\end{align}
 
-This paper presents the most thorough study to date of vehicular carrier-phase
-differential GNSS (CDGNSS) positioning performance in a deep urban setting
-unaided by complementary sensors.  Using data captured during approximately 2
-hours of driving in and around the dense urban center of Austin, TX, a CDGNSS
-system is demonstrated to achieve 17-cm-accurate 3D urban positioning (95\%
-probability) with solution availability greater than 87\%.  The results are
-achieved without any aiding by inertial, electro-optical, or odometry sensors.
-Development and evaluation of the unaided GNSS-based precise positioning
-system is a key milestone toward the overall goal of combining precise GNSS,
-vision, radar, and inertial sensing for all-weather high-integrity
-high-absolute-accuracy positioning for automated and connected vehicles.  The
-system described and evaluated herein is composed of a densely-spaced
-reference network, a software-defined GNSS receiver, and a real-time kinematic
-(RTK) positioning engine.  A performance sensitivity analysis reveals that
-navigation data wipeoff for fully-modulated GNSS signals and a dense reference
-network are key to high-performance urban RTK positioning.  A comparison with
-existing unaided systems for urban GNSS processing indicates that the proposed
-system has significantly greater availability or accuracy.
-\end{abstract}
+where $A_e = \frac{\lambda}{4 \pi}$ for an omnidirectional antenna and a signal
+with a wavelength $\lambda$. Substituting $A_e$ in (\ref{eq:general_PrPt})
+provides the Friis Formula, which applies for a wireless channel in a vacuum.
+Further adding terms $G_t$ and $G_r$ for the transmit and receive antenna gains,
+we reach the generalized Friis formula.
 
-\begin{IEEEkeywords} 
-urban vehicular positioning; CDGNSS; low-cost RTK positioning.
-\end{IEEEkeywords}
+\begin{align}
+  \label{eq:friss_formula}
+  P_r = \frac{G_t G_r \lambda ^2 P_t}{(4\pi d)^2}
+\end{align}
 
-%%% -------------------------- Select preprint or "for submission" version
-\newif\ifpreprint
-\preprintfalse
-%\preprinttrue
+\section{Path Loss Exponent Model}
+Since wireless channels exist in environments without a vacuum, a generalized
+model is needed to cover a variety of cases. A generalized such model is an
+exponential model using the received power formula below.
 
-\ifpreprint
+\begin{align}
+  \label{eq:general_exp_PL}
+  P_r = P_t G_t G_r K \left[\frac{d_0}{d}\right]^\alpha
+\end{align}
 
-\pagestyle{plain}
-\thispagestyle{fancy}  
-\fancyhf{} 
-\renewcommand{\headrulewidth}{0pt}
-\rfoot{\footnotesize \bf October 2020 preprint of paper submitted for review} \lfoot{\footnotesize \bf
-  Copyright \copyright~2020 by Lakhsay Narula \\ and Todd E. Humphreys}
+The parameter $d_0$ is a unit normalization parameter, usually set to $1$ m. The
+parameter $\alpha$ is the \gls{PL} exponent. The parameter $K$ is the nominal
+\gls{PL} at $1$ m. Notice when $K = [\frac{\lambda}{4 \pi}]^2$ defined as $K_0$
+and $\alpha = 2$, the formula reveals the Friis Formula from
+(\ref{eq:friss_formula}).
 
-\else
+\chapter{Multipath Effect Modeling}
+Other than the natural landscape of the environment, there are other effects
+that may further increase \gls{PL}.
 
-\thispagestyle{empty}
-\pagestyle{empty}
+\section{Blocking}
+Blocking allows for a multi-slope \gls{PL} model. Essentially, the \gls{PL}
+model has multiple definitions, and follows each one with a certain probability.
+This usually translates to following a certain \gls{PL} for a \gls{LOS} path,
+and following another model for the \gls{NLOS} path. A common model for this is
+the Exponential Blocking model
 
-\fi
+\begin{align}
+  \label{eq:exp_blocking}
+  P_r = \left\{ \begin{array}{ll}
+                  PL_{LOS}  & \text{w.p.} ~~~~~~ e^{-d/\beta}  \\
+                  PL_{NLOS} & \text{w.p.} ~~~ 1 - e^{-d/\beta} \\
+                \end{array}
+  \right.
+\end{align}
 
-%%% ----------------------------------------------------------------------
+where $\beta$ acts as the mean distance before blocking occurs (i.e. by a building), and the $PL_{LOS}$ and $PL_{NLOS}$ terns are functions for \gls{PL} models.
 
-\section{Introduction}
-In the introduction:
-\begin{itemize}
-\item {\bf Background:} Set the stage in a compelling way to draw the reader
-  in.
-\item {\bf Opportunity/Need:} Identify a need or an opportunity.
-\item {\bf The status quo:} Review the state of the art.  
-\item {\bf Gap:} Mind the gap!  Convince your reader that there is an
-  important deficiency (gap) in the existing literature or practice, and that
-  your paper will save the day by filling in/bridging this gap.  Otherwise,
-  why should your reader take time to read the paper?
-\item {\bf Vision:} Invite the reader to imagine the way the world could be if
-  only the gap could be bridged.
-\item {\bf Contributions:} Clearly enumerate the paper's contributions.
-\end{itemize}
+\subsection*{Blocking example}
+Imagine a scenario in which a \gls{LOS} path is in free space and the \gls{NLOS} is modeled as an exponential \gls{PL} with an exponent $\alpha = 2.5$ and a distance of $d = 100$ between the transmitter and receiver.
+The blocking exponent $\beta = 25$.
+What does the $SNR = P_r/P_t$ model finalize to if $G = G_t G_r = 1$, $K_{\text{LOS}} = K_{\text{NLOS}} = K_0 = -40$ dB and $d_0 = d_1 = 1$.
+\\
+\\
+Answer:
 
-What follows is an example introduction taken from
-\cite{humphreys2019deepUrbanIts}. \\ \\
+\begin{align}
+  SNR = \frac{P_r}{P_t} = \left\{ \begin{array}{ll}
+                                    K_0 \left[\frac{1}{100}\right]^2     & \text{w.p.} ~~~~~~ e^{-100/25}  \\
+                                    K_0 \left[\frac{1}{100}\right]^{2.5} & \text{w.p.} ~~~ 1 - e^{-100/25} \\
+                                  \end{array}
+  \right. \nonumber
+  = \left\{ \begin{array}{ll}
+              -80 \text{ dB} & \text{w.p.} ~ 0.0183 \\
+              -90 \text{ dB} & \text{w.p.} ~ 0.9817 \\
+            \end{array}
+  \right. \nonumber
+\end{align}
 
-\IEEEPARstart{F}{uture} Vehicle-to-vehicle (V2V) and vehicle-to-infrastructure
-(V2I) connectivity will permit vehicles to relay their positions and
-velocities to each other with millisecond latency, enabling tight coordinated
-platooning and efficient intersection management.  More ambitiously, broadband
-V2V and V2I enabled by 5G wireless networks will permit vehicles to share
-unprocessed or lightly-processed sensor data. \emph{Ad hoc} networks of
-vehicles and infrastructure will then function as a single sensing organism.
-The risk of collisions, especially with pedestrians and cyclists---notoriously
-unpredictable and much harder to sense reliably than vehicles---will be
-significantly reduced as vehicles and infrastructure contribute sensor data
-from multiple vantage points to build a blind-spot-free model of their
-surroundings.
+\section{Shadowing}
+Shadowing is attenuation caused by objects between the transmitter and
+receiver that reduce the received signal's power.
+It is modeled as random and does not depend on the distance of the wireless communication link.
+By modeling shadowing we account for the constructive and destructive interference cause by the transmitted signal reflecting, diffracting and scattering, the components of which are called multipath components.
 
-Such collaborative sensing and traffic coordination requires vehicles to know
-and share their own position.  How accurately?  The proposed Dedicated Short
-Range Communications (DSRC) basic safety message, a first step in V2V
-coordination, does not yet define a position accuracy requirement, effectively
-accepting whatever accuracy a standard GNSS receiver provides
-\cite{kenney2011dedicated}.  But automated intersection management
-\cite{fajardo2011automated}, tight-formation platooning, and unified
-processing of sensor data---all involving vehicles of different makes that may
-not share a common map---will be greatly facilitated by globally-referenced
-positioning with sub-30-cm accuracy.
+The shadowing effect is modeled as lognormal.
+The received power equation from a \gls{PL} model is simply multiplied with a new variable $\mathcal{X}$, where $x_{\text{dB}} = 10 log_{10}{\mathcal{X}} \sim \mathcal{N}(0,\sigma_{\text{dB}}^2)$.
 
-Poor weather also motivates high-accuracy absolute positioning.  Every
-automated vehicle initiative of which the present authors are aware depends
-crucially on lidar or cameras for fine-grained positioning within their local
-environment.  But these sensing modalities perform poorly in low-visibility
-conditions such as a snowy whiteout, dense fog, or heavy rain.  Moreover,
-high-definition 3D maps created with lidar and camera data, maps that have
-proven crucial to recent progress in reliable vehicle automation, can be
-rendered dangerously obsolete by a single snowstorm, leaving vehicles who rely
-on such maps for positioning no option but to fall back on GNSS and radar to
-navigate a snow-covered roadway in low-visibility conditions.  When, as is
-often the case on rural roads, such snowy surroundings offer few
-radar-reflective landmarks, radar too becomes useless.  GNSS receivers operate
-well in all weather conditions, but only a highly accurate GNSS solution,
-e.g., one whose absolute errors remain under 30 cm 95\% of the time, could
-prevent a vehicle's drifting onto a snow-covered road's soft shoulder.  Code-
-and Doppler-based GNSS solutions can be asymptotically accurate (averaged over
-many sessions) to better than 50 cm, which may be adequate for digital mapping
-\cite{narula2018accurate}, but they will find it challenging to meet a 30 cm
-95\% stand-alone requirement, even with modernized GNSS offering wideband
-signals at multiple frequencies.
+Alternatively, some use a model where $P_r = P_t \Psi$ where $ \psi_{\text{dB}} = 10 log_{10}{\Psi}  \sim \mathcal{N}(PL_{\text{dB}}(d),\sigma_{\text{dB}}^2)$ for some \gls{PL} model.
 
-Carrier-phase-based GNSS positioning---also referred to as precise GNSS
-positioning even though it actually offers absolute accuracy, not just
-precision (repeatability)---can meet the most demanding accuracy requirements
-envisioned for automated and connected vehicles, but has historically been
-either too expensive or too fragile, except in open areas with a clear view of
-the overhead satellites, for widespread adoption.  Coupling a carrier-phase
-differential GNSS (CDGNSS) receiver with a tactical grade inertial sensor, as
-in \cite{petovello2004benefits,scherzinger2006precise,
-  zhangComparisonWithTactical2006,kennedy2006architecture} enables robust
-high-accuracy positioning even during the extended signal outages common in
-dense urban areas.  But GNSS-inertial systems with tactical-grade inertial
-measurement units (IMUs) cost tens of thousands of dollars and have proven
-stubbornly resistant to commoditization.  Coupling a GNSS receiver with
-automotive- or industrial-grade IMUs is much more economical, and
-significantly improves performance, as shown in \cite{li2018high}.  But such
-coupling only allows approximately 5 seconds of complete GNSS signal blockage
-before the IMU no longer offers a useful constraint for so-called integer
-ambiguity resolution \cite{evaluationLowCostMems2006Godha}, which underpins
-the fastest, most accurate, and most robust CDGNSS techniques, namely,
-single-baseline RTK, network RTK, and PPP-RTK
-\cite{teunissen2015review,cui2017rtkforCAV}.
+\subsection*{Shadowing example}
+Imagine a scenario in which we wish to evaluate Wi-Fi coverage for a desired range of $100$ m, with a minimum $SNR = 5$ dB and outage constraint at $1\%$ for the $5$ GHz frequencies.
+The \gls{PL} is modeled as exponential with $\alpha = 3$, $K = K_0 = -50$ dB (for $F_c \approx 5$ GHz), no TX and RX gain, a noise power of $P_n = -100$ dBm and a shadowing standard deviation of $\sigma_{\text{dB}} = 6$.
+What is the mimum transmition power to meet the outage constraint?
+\\
+\\
+Answer:
 
-Previous research has suggested an inexpensive technique for robustifying RTK
-positioning: tightly coupling carrier-phase-based GNSS positioning with
-inertial sensing and vision \cite{shepard2014fusion,pesyna2015dissertation}.
-Such coupling takes advantage of the remarkable progress in high-resolution,
-low-cost cameras within the intensely competitive smartphone market.  The
-current authors are engaged in developing a high-integrity RTK-vision system
-for high-accuracy vehicular positioning in rural and urban environments.
-Further coupling with radar will make the system robust to low-visibility
-conditions.
+\begin{align}
+  SNR             & = \frac{P_r}{P_t} \geq 5 \text{ dB} \nonumber                                       \\
+  P_r             & \geq -95 \text{ dB} \nonumber                                                       \\
+  P_r             & = P_{t,\text{dB}} + K_{0,\text{dB}} + x_{\text{dB}} - \alpha 10 log_{10}d \nonumber \\
+  P_{t,\text{dB}} & \geq -95 + 50 - x_{\text{dB}} + 60 \nonumber                                        \\
+  P_{t,\text{dB}} & \geq 15 - x_{\text{dB}} \nonumber
+\end{align}
 
-As a step toward this goal, it is of interest to evaluate the performance of
-stand-alone RTK techniques---those unaided by IMUs, odometry, or vision---in
-urban environments.  Such a study will reveal why and when aiding is
-necessary, and how an RTK positioning system might behave if aiding were
-somehow impaired or unavailable, whether due to sensor faults or, in the case
-of exclusive visual aiding, poor visibility conditions.
+Since $x_{\text{dB}}$ is normally distributed, we simply use the CDF of a Gaussian distribution to evaluate transmit power that would allow for the required outage constraint of $0.01$:
 
-Little prior work has explored unaided vehicular RTK performance in urban
-environments, no doubt because performance results have historically been
-dismal. Short-baseline RTK experiments between two vehicles in
-\cite{ong2009assessment} revealed that multi-frequency (L1-L2) GPS and Glonass
-RTK yielded poor results in residential and urban environments.  Only along a
-mountain highway with a relatively clear view of the sky was availability
-greater than 90\% and accuracy better than 30 cm.  RTK positioning in downtown
-Calgary was disastrous, with less than 60\% solution availability and RMS
-errors exceeding 9 meters.
+\begin{align}
+  P(P_{t,\text{dB}} \geq  15 - x_{\text{dB}}) = 1 - 0.01 \nonumber \\
+  15 - P_{t,\text{dB}} = \sigma_{\text{dB}}Q^{-1}(0.99) \nonumber  \\
+  P_{t,\text{dB}} \geq 29 \text{ dBm} \nonumber
+\end{align}
 
-More recently, Li et al. \cite{li2018high} have shown that, with the benefit
-of greater signal availability, unaided professional-grade dual-frequency GPS
-+ BDS + GLONASS RTK can achieve correct integer fixing rates of 76.7\% on a
-1-hour drive along an urban route in Wuhan, China.  But Li et al. do not
-provide data on the incorrect fixing rate, nor a full error distribution, so
-the significance of their results is difficult to assess.
+\chapter{Multipath Channel Models}
 
-Recent urban RTK testing by Jackson et al. \cite{jackson2018assessmentRtk}
-indicates that no low-to-mid-range consumer RTK solution offers greater than
-35\% fixed (integer-resolved) solution availability in urban areas, despite a
-dense reference network and dual-frequency capability.  A key failing of
-existing receivers appears to be their slow recovery after passing under
-bridges or overpasses.
+\glsaddall
+\printglossary[type=\acronymtype,title=Acronyms]
 
-This paper describes and evaluates an unaided RTK positioning system that has
-been designed for vehicular operation in both rural and urban environments.
-Preliminary performance results were published in a conference version of this
-paper \cite{humphreys2018urbanStandAloneRTKplans2018}.  The current paper
-improves on the conference version in four ways: (1) the test route is both
-more challenging and more comprehensive, (2) a proper independent ground truth
-trajectory is used as the basis of error evaluation, (3) data modulation
-wipeoff for improved carrier tracking robustness is applied not only on GPS L1
-C/A signals, as previously, but now also on SBAS L1 signals, and (4) the
-performance benefit of vehicle GNSS antenna calibration is assessed.
-
-This paper's primary contributions are (i) a demonstration of the performance
-that can be achieved with a low-cost software-defined unaided RTK GNSS
-platform in a dense urban environment, and (ii) an evaluation of the relative
-importance of various factors (e.g., data bit wipeoff, age of reference data,
-rover antenna calibration, reference network density) to the overall system
-performance.
-
-To stimulate further innovation in urban precise positioning, all data from
-this paper's urban driving campaign have been posted at
-\url{http://radionavlab.ae.utexas.edu} under ``Public Datasets,'' including
-wideband (10 MHz) intermediate frequency samples from both the reference and
-rover antennas, RINEX-formatted rover and reference observables, and the
-ground truth trajectory.
-
-\section{Tables}
-Captions should appear above tables.  Use the full word Table to refer to a
-table in your paper; e.g., ``Results of the study are shown in Table I.''
-Follow the format of Table \ref{tab:Chi2Fits} and Table
-\ref{tab:ValidationTable2}.
-
-\begin{table}[ht]
-  \centering
-  \caption{Chi-square Values for Fits to Nakagami-m and Rice Distributions}
-  \begin{tabular}[c]{cccc}
-    \toprule
-    Data Source &  Sets, DOF & Nakagami-m & Rice \\ \midrule
-    Wideband UHF & 79, ~~8 & 11.8 $\pm$ 8.8 & 9.0 $\pm$ 4.3 \\
-    GPS L$_1$ & 33, ~~7 & 8.42 $\pm$ 5.9 & 7.7 $\pm$ 5.7 \\ \bottomrule
-  \end{tabular}
-  \label{tab:Chi2Fits}  
-\end{table}
-
-\begin{table}[ht]
-  \centering
-  \caption{Scintillation Effects Comparison:~~Empirical Truth Data}
-  \begin{tabular}[c]{ccccccc}
-    \toprule
-    \multicolumn{3}{c}{Parameters}  
-    & \multicolumn{2}{c}{Truth Scint.} 
-    & \multicolumn{2}{c}{Synthetic Scint.} \\
-    \cmidrule(r){1-3} \cmidrule(r){4-5} \cmidrule(r){6-7}
-    $S_4$& $\tau_0$ (s) & $T$ (s) & $\sigma_{\varphi}$ (deg) 
-    & $N_s$ & $\sigma_{\varphi}$ (deg) & $N_s$ \\  \midrule
-    0.87 & 0.18 & 200 & 16.4 & 32 & 17.5 $\pm$ 0.5 & 35.9 $\pm$ 4.7 \\ 
-    1.0 & 0.36 & 265 & 14.1 & 37 & 15.0 $\pm$ 0.5 & 41.6 $\pm$ 5.9 \\  
-    0.69 & 0.18 & 174 & 12.7 & 12 & 11.8 $\pm$ 0.9 & 5.6 $\pm$ 1.6 \\ 
-    0.87 & 0.26 & 225 & 11.6 & 23 & 12.7 $\pm$ 0.5 & 19.2 $\pm$ 4.6 \\
-    0.61 & 0.47 & 162 & 3.96 & 0 & 3.63 $\pm$ 0.2 & 0.10 $\pm$ 0.3 \\
-    0.96 & 0.09 & 81 & 28.5 & 60 & 32.7 $\pm$ 1.0 & 69.4 $\pm$ 5.8 \\
-    0.95 & 0.26 & 123 & 14.1 & 21 & 15.6 $\pm$ 0.5 & 19.8 $\pm$ 3.7 \\
-    0.51 & 0.71 & 138 & 2.12 & 0 & 1.60 $\pm$ 0.1 & 0 $\pm$ 0~~~ \\
-    \bottomrule
-  \end{tabular}
-  \label{tab:ValidationTable2}  
-\end{table}
-
-\section{Figures}
-\begin{enumerate}
-\item Use Fig. instead of Figure to refer to a figure (per IEEE style guide).
-  E.g., ``Fig. (4) illustrates this point by plotting $x$ vs. $y$.''
-\item Only use color in figures if necessary.  Many figures look better in
-  print if done in grayscale.  See the example figures in the writing templates.
-\item Try to make figure captions sufficient, or nearly so, to describe the
-  figure.  For example, the caption should explain what the various lines or
-  enclosed regions in a figure mean.  Push description from the text to the
-  caption to avoid duplication.
-\item Avoid legends in figures; prefer labels that point with lines directly
-  to the object being labeled.
-\item Make figures as small as they can be while remaining clearly legible and
-  conveying their message.  Often, you can use {\tt subplot(211)} in Matlab to
-  cut a figure height in half.
-\item Figures should not have titles on top.  All relevant information should
-  be put in the caption.
-\item Ensure that each figure's x and y axis labels, and all internal labels,
-  are in the same or similar font as the text and are large enough to read
-  easily.  A rule of thumb: make labels no smaller than the caption font.  You
-  can use {\tt figset.m} to change native Matlab font to one that more closely
-  resembles LaTeX default font and increase its size.
-\end{enumerate}
-
-Have a look at Fig. \ref{fig:rocplot1} as an example.  Todd uses an
-application called xfig to add labels and LaTeX symbols directly to eps
-figures.  It's old school, but he hasn't found a better alternative for
-perfectly matching the LaTeX fonts in the text.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width = 8.5cm]
-{figs/rocplot1}
-\caption{ROCs for the ML, MAP, and MMSE security code estimation strategies
-  under the following scenario: $(\cnr)_s = 54$ dB-Hz,
-  $(\cnr)_r = 48$ dB-Hz, $\eta = 1$, $d = 0$, $N = 400$. Top panel: High-rate
-  security code with $T_w = 2~ \mu$s.  Bottom panel: Low-rate security code
-  with $T_w = 20$ ms.}
-\label{fig:rocplot1}
-\end{figure}
-
-Sometimes you have to take up both columns with a figure, as in
-Fig. \ref{fig:rx-model}.
-
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=\textwidth]{figs/correlation}
-  \caption{Block diagram of the standard AGC, correlation, and accumulation
-    operations in a GNSS receiver. The product of the AGC-scaled incoming
-    signal $\beta(t)r(t)$ and the conjugate of the local replica
-    $\ell(t,\tau)$ is accumulated over $T$ seconds to produce the discrete
-    complex-valued accumulation product $\xi_k(\tau)$. For notational
-    convenience, the accumulation product has been scaled by $1/T$.}
-  \label{fig:rx-model}
-\end{figure*}
-
-\section{Algorithms}
-Use the full word Algorithm to refer to an algorithm in your paper; e.g.,
-``Algorithm~\ref{algo:correlation-search} shows the pseudocode for this
-process.'' Use the following format for algorithms. Note that variables should
-be written in mathematical italic font, whereas functions can be written either
-in typewriter font; e.g.,
-\[D_{\text m} = \texttt{toOGM} \left( \bm{p}_{\text m}^{W} - \bm{t}_k^{V},
-    \delta t \right)\] or in mathematical italic font.
-
-\begin{algorithm}
-  \footnotesize
-  \SetKwInOut{Input}{Input}
-  \SetKwInOut{Output}{Output}
-  \Input{$\bm{p}_{\text m}^W$, $\bm{p}_{{\text b, } 1:k}^B$, $\bm{t}_{1:k}^{V}$, $\phi_{1:k}^{V}$, $\sigma_t$, $\sigma_\phi$, $\delta t$, $\delta \phi$}
-  \vspace{0.25em}
-  \Output{$\widehat{\bm{\Theta}}$}
-
-  \vspace{1em}
-
-  $D_{\text m} = \texttt{toOGM} \left( \bm{p}_{\text m}^W - \bm{t}_k^{V}, \delta t \right)$
-
-  $\tilde{D}_{\text m} = \texttt{FFT2} \left( \texttt{pad} \left( D_{\text m}, 3\sigma_t \right) \right)$
-
-  \vspace{1em}
-
-  $\bm{p}_{{\text b, } 1:k}^{V} = R \left( \phi_{1:k}^{V} \right) \bm{p}_{{\text b, } 1:k}^B + \bm{t}_{1:k}^{V}$
-
-  $D_{\text b} = \texttt{toOGM} \left( \bm{p}_{{\text b, } 1:k}^{V} - \bm{t}_k^{V}, \delta t \right)$
-
-  $\tilde{D}_{\text b} = \texttt{FFT2} \left( \texttt{pad} \left( D_{\text b}, 3\sigma_t \right) \right)$
-
-  \vspace{1em}
-
-  $n = 3\sigma_t / \delta t$
-
-  $m = 3\sigma_\phi / \delta \phi$
-
-  \For{i = -m:m}{ 
-    $\Delta \phi = i \delta \phi$
-
-    $\tilde{D}_{\text b}^{\Delta \phi} = \texttt{rotate2} \left( \tilde{D}_{\text b}, \Delta \phi \right)$
-
-    $\mathsf{R}[i,:,:] = \texttt{IFFT2} \left( \tilde{D}_{\text m} \circ \texttt{conj} \left( \tilde{D}_{\text b}^{\Delta \phi} \right) \right) \left[ -n : n, -n : n \right]$
-  }
-  $\widehat{\bm{\Theta}} = \argmax \left( \mathsf{R} \right)$
-
-  \caption{\texttt{fastGlobalAlign}}
-  \label{algo:correlation-search}
-\end{algorithm}
-
-\begin{algorithm}
-\SetKwInOut{Input}{Input}
-\SetKwInOut{Output}{Output}
-\Input{$\vbh{z} \in \mathbb{R}^m$, $L \in \mathbb{R}^{m \times m}$}
-\Output{$\vbc{z} \in \mathbb{Z}^m$}
-$\vbh{z}_{\text c} = \vbh{z}$
-
-\For{i = 1:m}{ 
-$\check{{z}}_{i}=\lfloor\hat{{z}}_{{\text c}i} \rceil$
-
-$\check{{\epsilon}}_{{\text c}i}=\hat{{z}}_{{\text c}i} - \check{{z}}_{i}$
-
-\For{j = i+1:m}{
-$\hat{{z}}_{{\text c}j}=\hat{{z}}_{{\text c}j} - l_{ij}\check{{\epsilon}}_{{\text c}i} $\
-}
-}
-\caption{${\text{IB}}( \vbh{z}, L)$}
-\label{algo:second-algo}
-\end{algorithm}
-\noindent
-
-\section{Stylistic Conventions}
-\subsection{Mathematical Notation}
-\begin{enumerate}
-\item Define \emph{all} mathematical notation that you introduce unless it's
-  been previously defined.  Where possible, introduce notation \emph{before}
-  presenting the equation involving the notation.
-\item Put units in straight font, not italics or math font; e.g., cm not $cm$.
-\item Write subscripts that represent indices or are just one letter long in
-  the usual mathematical italic font; e.g., ${x}_{ak}$.  Write subscripts that
-  are words or abbreviated words in roman type: ${x}_{\text{max}}$.
-\item Defining {\tt newcommand} macros for the paper can make it easier to
-  repeatedly write complicated LaTeX math expressions; e.g., $\Fc{1}$, $\Fs$,
-  $\Fsr$.
-\end{enumerate}
-
-\subsection{Punctuation}
-Use American Style punctuation.
-\href{https://www.thepunctuationguide.com/british-versus-american-style.html}{This}
-website is a good reference. Note that American style quotations include
-commas and periods within the quotation, even if they are not actually part of
-what was quoted; e.g.,
-\begin{quotation}
-  “Economic systems,” according to Professor White, “are an inevitable
-  byproduct of civilization, and are, as John Doe said, ‘with us whether we
-  want them or not.’”
-\end{quotation}
-This style is not as sensible as British English, but it's the convention on
-this side of the pond.
-
-\subsection{Abbreviations}
-\begin{enumerate}
-\item With few exceptions, define all abbreviations/acronyms used in
-  the paper (e.g., UAV, GNSS, IMU, etc.).  Exceptions are those that have
-  become so common that they're part of everyday vernacular (radar, GPS). 
-\end{enumerate}
-
-\subsection{References}
-\begin{enumerate}
-\item Never start a sentence with the actual numerical citation (e.g., ``[2]
-  presents an optimal ...'').  Prefer instead ``Reference [2] ...'' or ``The
-  authors of [2]...'' or ``Durge et al. present in [2] an optimal ...''
-\end{enumerate}
-
-\subsection{Signposts}
-Busy readers rarely read a paper straight through from abstract to conclusions.
-Instead, they tend to sample bits and pieces \emph{a la carte}.  For these
-readers, and even for sequential readers who may need some guidance on where
-they are and what is to come, it's helpful to start off each major section with
-a \emph{signpost}---a brief summary of what the reader now knows and what he
-can expect to learn from that section.  Sometimes the section title, perhaps in
-combination with the first sub-section title, sufficies.  In other cases, a few
-sentences are needed to orient the reader.  Here is an example of a
-few-sentence signpost at the beginning of a section:
-\begin{quotation}
-  The model of road user behavior presented above applies to human
-  decisions. Autonomous vehicles are likely to respond very differently,
-  because they are fundamentally unable to win the game of chicken. Moreover,
-  they are likely to obey the rules of the road far more than human drivers,
-  particularly in safety-critical situations. This section elaborates such
-  reasoning.
-\end{quotation}
-
-\section{Conclusions}
-Like the abstract, this section should give a recap of the whole paper.  Write
-in simple past tense.  Don't assume the reader has read the paper.
-
-Example:\\
-
-A real-time kinematic (RTK) positioning system tailored for urban vehicular
-positioning has been described and evaluated.  To facilitate performance
-comparison against similar systems, the system was tested without any benefit
-of aiding by inertial or electro-optical sensors.  Over nearly 2 hours of
-urban testing, including multiple passes through Austin's dense urban center,
-the system achieved an 85\% probability of correct integer fix for a 2.4\%
-probability of incorrect fix, resulting in 3D positioning errors smaller than
-17 cm (95\%).  A performance sensitivity analysis revealed that navigation
-data bit prediction on fully-modulated GNSS signals is key to high-performance
-urban RTK positioning, and that a dense reference network, carrier tracking
-bandwidth adaptation, and rover antenna calibration each offer a significant
-integrity benefit. A comparison with existing unaided systems for urban GNSS
-processing indicates that the proposed system has a significant advantage in
-availability and/or accuracy.
-
-\section*{Acknowledgments}
-Example:\\
-
-This work was supported in part by the U.S. Department of Transportation (USDOT)
-under the University Transportation Center (UTC) Program Grant 69A3552047138
-(CARMEN), and by affiliates of the 6G@UT center within the Wireless Networking
-and Communications Group at The University of Texas at Austin.
-
-\bibliographystyle{ieeetran} 
-\bibliography{pangea}
+% \bibliographystyle{ieeetran}
+% \bibliography{ref}
 \end{document}

### 2023-02-08 paper.tex
SNR and constellation wip

@@
 \newacronym{NLOS}{NLOS}{Non-line of sight}
+\newacronym{BW}{BW}{Bandwidth}
+\newacronym{iid}{i.i.d.}{independent and identically-distributed}
+\newacronym{CLT}{CLT}{Central Limit Theorem}
+\newacronym{GRV}{GRV}{Gaussian random variable}
+\newacronym{SNR}{SNR}{signal-to-noise ratio}
+\newacronym{PSD}{PSD}{power spectral density}
+\glsaddall
 
@@
 \subsection*{Blocking example}
-Imagine a scenario in which a \gls{LOS} path is in free space and the \gls{NLOS} is modeled as an exponential \gls{PL} with an exponent $\alpha = 2.5$ and a distance of $d = 100$ between the transmitter and receiver.
-The blocking exponent $\beta = 25$.
-What does the $SNR = P_r/P_t$ model finalize to if $G = G_t G_r = 1$, $K_{\text{LOS}} = K_{\text{NLOS}} = K_0 = -40$ dB and $d_0 = d_1 = 1$.
+Imagine a scenario in which a \gls{LOS} path is in free space and the \gls{NLOS}
+is modeled as an exponential \gls{PL} with an exponent $\alpha = 2.5$ and a
+distance of $d = 100$ between the transmitter and receiver. The blocking
+exponent $\beta = 25$. What does the $SNR = P_r/P_t$ model finalize to if $G =
+  G_t G_r = 1$, $K_{\text{LOS}} = K_{\text{NLOS}} = K_0 = -40$ dB and $d_0 = d_1 =
+  1$.
 \\
@@
 \section{Shadowing}
-Shadowing is attenuation caused by objects between the transmitter and
-receiver that reduce the received signal's power.
-It is modeled as random and does not depend on the distance of the wireless communication link.
-By modeling shadowing we account for the constructive and destructive interference cause by the transmitted signal reflecting, diffracting and scattering, the components of which are called multipath components.
+Shadowing is attenuation caused by objects between the transmitter and receiver
+that reduce the received signal's power. It is modeled as random and does not
+depend on the distance of the wireless communication link. By modeling shadowing
+we account for the constructive and destructive interference cause by the
+transmitted signal reflecting, diffracting and scattering, the components of
+which are called multipath components.
 
-The shadowing effect is modeled as lognormal.
-The received power equation from a \gls{PL} model is simply multiplied with a new variable $\mathcal{X}$, where $x_{\text{dB}} = 10 log_{10}{\mathcal{X}} \sim \mathcal{N}(0,\sigma_{\text{dB}}^2)$.
+The shadowing effect is modeled as lognormal. The received power equation from a
+\gls{PL} model is simply multiplied with a new variable $\mathcal{X}$, where
+$x_{\text{dB}} = 10 log_{10}{\mathcal{X}} \sim
+  \mathcal{N}(0,\sigma_{\text{dB}}^2)$.
 
-Alternatively, some use a model where $P_r = P_t \Psi$ where $ \psi_{\text{dB}} = 10 log_{10}{\Psi}  \sim \mathcal{N}(PL_{\text{dB}}(d),\sigma_{\text{dB}}^2)$ for some \gls{PL} model.
+Alternatively, some use a model where $P_r = P_t \Psi$ where $ \psi_{\text{dB}}
+  = 10 log_{10}{\Psi}  \sim \mathcal{N}(PL_{\text{dB}}(d),\sigma_{\text{dB}}^2)$
+for some \gls{PL} model.
 
 \subsection*{Shadowing example}
-Imagine a scenario in which we wish to evaluate Wi-Fi coverage for a desired range of $100$ m, with a minimum $SNR = 5$ dB and outage constraint at $1\%$ for the $5$ GHz frequencies.
-The \gls{PL} is modeled as exponential with $\alpha = 3$, $K = K_0 = -50$ dB (for $F_c \approx 5$ GHz), no TX and RX gain, a noise power of $P_n = -100$ dBm and a shadowing standard deviation of $\sigma_{\text{dB}} = 6$.
-What is the mimum transmition power to meet the outage constraint?
+Imagine a scenario in which we wish to evaluate Wi-Fi coverage for a desired
+range of $100$ m, with a minimum $SNR = 5$ dB and outage constraint at $1\%$ for
+the $5$ GHz frequencies. The \gls{PL} is modeled as exponential with $\alpha =
+  3$, $K = K_0 = -50$ dB (for $F_c \approx 5$ GHz), no TX and RX gain, a noise
+power of $P_n = -100$ dBm and a shadowing standard deviation of
+$\sigma_{\text{dB}} = 6$. What is the mimum transmition power to meet the outage
+constraint?
 \\
@@
 
-Since $x_{\text{dB}}$ is normally distributed, we simply use the CDF of a Gaussian distribution to evaluate transmit power that would allow for the required outage constraint of $0.01$:
+Since $x_{\text{dB}}$ is normally distributed, we simply use the CDF of a
+Gaussian distribution to evaluate transmit power that would allow for the
+required outage constraint of $0.01$:
 
@@
 \chapter{Multipath Channel Models}
+\label{multipath_ch_model}
+A transmitted signal $s(t)$ propagates through an environment in multiple paths.
+Neglecting noise, the received signal is simply a sum of attenuated, delayed,
+and phase shifted versions of the original transmitted signal. The different
+paths a signal takes to get to a receiver results in a sum with patterns of high
+or low destructive interference, an effect called multipath fading.
+\section{Multipath Model}
+We model the channel $c(\tau,t)$ as a time-varying ($t$) impulse response function evaluated at $t - \tau$.
+As with any system, the received signal is modeled as a convolution of $c(\tau,t)$ with the transmitted signal $s(t)$.
 
-\glsaddall
+A way to interpret the parameters is as follows:
+\begin{itemize}
+  \item $t$ is the time an impulse response is observed
+  \item $\tau$ is the offset from the observation time in which the impulse was transmitted
+\end{itemize}
+
+The traditional impulse response of a time-invariant system can be expressed as
+thinking of the channel's impulse response as periodic $c(\tau,t + T)$ where $T
+  = -t$ such that  $c(\tau,0) = c(\tau)$, which is the impulse response at time
+$0$.
+
+Whether \gls{NLOS} received signals are received in a large spread of time
+compared to the \gls{LOS} component or not, depends on the inverse of the
+transmitted signal's \gls{BW}. This spread is called the delay spread $T_m$.
+Depending on whether it is large or small changes the model significantly, so it
+is explored as Narrowband and Wideband fading.
+
+\section{Narrowband Fading Model}
+Under this regime, the delay spread $T_m$ is significantly smaller than the inverse of the baseband \gls{BW} of the signal $B_{bb}$, $T_m << B^{-1}_{bb}$.
+Essentially, this model assumes that all paths have no delay as mentioned in Section \ref{multipath_ch_model}, but still have random phases.
+For a received signal $r(t)$, the following model is assumed,
+
+\begin{align}
+  r(t) = \operatorname{Re} \Biggl\{ s(t) e^{j2\pi f_c t} \left(\sum_i \alpha_i(t) e^{-j \phi_i(t)}\right) \Biggr\}
+\end{align}
+where $f_c$ is the carrier frequency, $\alpha_i$ are the attenuation on each path $i$ with their respective phase offset $\phi_i$.
+
+For large $N(t)$ (time dependent number of multipath components) if $\alpha_i$ and $\phi_i$ are \gls{iid}, then by the \gls{CLT}, the real and imaginary components of the baseband equivalent received signal are approximately joint \glspl{GRV}.
+If some path is larger in amplitude, like the \gls{LOS} path, then the paths are no longer \gls{iid} so the \gls{CLT} no longer holds and the baseband components are not joint \gls{GRV}.
+
+TODO Autocorrelation
+\chapter{Channel Reliability}
+
+\section{Signal Modulation}
+
+TODO
+
+\begin{figure*}[t]
+  \centering
+  \includegraphics[width=0.75\textwidth]{figs/constellations.pdf}
+  \caption{Constellation schemes where Quadrature and In-Phase possible discrete options are displayed in terms of their energy.}
+  \label{fig:constellations}
+\end{figure*}
+
+\begin{table}[htbp]
+  \centering
+  \caption{}
+  \begin{tabular}{lcccc}                                                                                                                   \\
+    \cmidrule(r){1-5}  & MPAM                           & MPSK                       & MQAM                         & FSK     \\
+    \midrule $E_s$     & $\frac{d_{min}^2}{12}(M^2-1)$  & $d^2$                      & $\frac{d_{min}^2}{6}(M^2-1)$ & $d^2/2$ \\
+    \midrule $d_{min}$ & $\sqrt{\frac{12E_s}{M^2 - 1}}$
+                       & $\sqrt{E_s}$                   & $\sqrt{\frac{6}{M^2 - 1}}$ & $\sqrt{2 E_s}$                         \\
+
+    \bottomrule
+  \end{tabular}
+  \label{tab:win-space-results}
+\end{table}
+
+\section{Signal-to-noise ratio}
+
+If the complex noise $n(t) \sim \mathcal{C}\mathcal{N}(0,2\sigma^2)$, meaning
+each of the two channels is distributed as a zero-mean \gls{GRV} with variance
+$\sigma$, this relates to the noise \gls{PSD} $N_0$ (in watts per hertz $W/Hz$)
+as $\sigma^2 = N_0/2$. If the energy per dimension is $\bar{E_s} = E_s/2$, then
+the \gls{SNR} is defined as:
+
+\begin{align}
+  \gamma_s = SNR = \frac{\bar{E_s}}{\sigma} = \frac{E_s}{2\sigma} = \frac{E_s}{N_0}
+\end{align}
+
+\Gls{SNR} can also be defined in terms of the power of the received signal $P_s$
+over the power of the noise $P_n = N_0 B$, where $B = 1/T$ is the \gls{BW} of
+the receiver. Note, the received signal power $P_s$ relates to the energy per
+symbol as $P_s = E_s/T$ for a symbol time T.
+
+\begin{align}
+  \gamma_s = SNR = \frac{P_s}{P_n}= \frac{E_s/T}{N_0 B} = \frac{E_s}{N_0}
+\end{align}
+
+The \gls{SNR} also has an equivalent metric, the per-bit \gls{SNR} $Ebb-N_0$
+expressed in terms of the energy per bit $E_b$ over the noise \gls{PSD} $N_0$.
+The energy per bit is $E_b = E_s/\log _{2} M$. Also, define the bit-rate as $R =
+  \frac{\log _{2} M}{T}$.
+
+\begin{align}
+  \gamma_b = Ebb-N_0 = \frac{E_b}{N_0} = \frac{ E_s/\log _{2} M}{N_0} = \frac{\gamma_s}{\log _{2} M} = \gamma_s \frac{B}{R}
+\end{align}
+
+\section{Probability of Error}
+TODO
+
+The probability of error $P_s$ is defined in a per-symbol basis. Each constellation has a different $P_s$, seen in the table below based on the Nearest Neighbor bound.
+
+\begin{table}[htbp]
+  \centering
+  \caption{}
+  \begin{tabular}{lcccc}                                                                                \\
+    \cmidrule(r){1-3} & $P_s(\gamma_s)$                                  & $P_b(\gamma_b)$ \\
+    \midrule MPAM     & $2(1-1/M)Q(\sqrt{\frac{3\gamma_s}{M^2-1}})$      &                 \\
+    \midrule MPSQ     & $2Q(\sqrt{2\gamma_s}\sin(\pi/M))$                                  \\
+    \midrule MQAM     & $4(1-1/\sqrt{M})Q(\sqrt{\frac{3\gamma_s}{M-1}})$                   \\
+    \midrule FSK      & $(M-1)Q(\sqrt(\gamma_s/2))$                                        \\
+    \\
+
+    \bottomrule
+  \end{tabular}
+  \label{tab:constellation_pe}
+\end{table}
+
+Useful approximations for $0 \leq \gamma_s \leq 30$ dB:
+\begin{itemize}
+  \item MQAM $M \geq 4$ has $P_b(\gamma_s) \approxeq 0.2 e^{\frac{-1.5\gamma_s}{M-1}}$
+  \item BPSK has $P_b(\gamma_s) = P_s(\gamma_s) \approxeq 2 e^{-1.5\gamma_s}$
+\end{itemize}
+
 \printglossary[type=\acronymtype,title=Acronyms]

### 2023-02-09 paper.tex
1/4gamma avg BPSK found

@@
  \makeglossaries
+ \usepackage{mdframed}
 
@@
 \newacronym{PSD}{PSD}{power spectral density}
+\newacronym{BER}{BER}{bit error rate}
+\newacronym{AWGN}{AWGN}{additive white Gaussian noise}
+
 \glsaddall
@@
 where $\beta$ acts as the mean distance before blocking occurs (i.e. by a building), and the $PL_{LOS}$ and $PL_{NLOS}$ terns are functions for \gls{PL} models.
+\begin{mdframed}
 
@@
   \end{align}
+\end{mdframed}
 
@@
 
+\begin{mdframed}
   \subsection*{Shadowing example}
@@
   \end{align}
-
+\end{mdframed}
 \chapter{Multipath Channel Models}
@@
 
-\section{Probability of Error}
+\section{Probability of Error for fixed SNR}
 TODO
 
-The probability of error $P_s$ is defined in a per-symbol basis. Each constellation has a different $P_s$, seen in the table below based on the Nearest Neighbor bound.
+The probability of error $P_s$ is defined in a per-symbol basis. Each constellation has a different $P_s$, seen in the table below based on the Nearest Neighbor bound in an \gls{AWGN} channel.
 
@@
     \cmidrule(r){1-3} & $P_s(\gamma_s)$                                  & $P_b(\gamma_b)$                                                             \\
-    \midrule MPAM     & $2(1-1/M)Q(\sqrt{\frac{3\gamma_s}{M^2-1}})$      &                 \\
-    \midrule MPSQ     & $2Q(\sqrt{2\gamma_s}\sin(\pi/M))$                                  \\
-    \midrule MQAM     & $4(1-1/\sqrt{M})Q(\sqrt{\frac{3\gamma_s}{M-1}})$                   \\
-    \midrule FSK      & $(M-1)Q(\sqrt(\gamma_s/2))$                                        \\
-    \\
-
+    \midrule MPAM     & $2(1-1/M)Q(\sqrt{\frac{6\gamma_s}{M^2-1}})$      & $\frac{2(1-1/M)}{\log_{2}M}Q(\sqrt{\frac{6\gamma_b\log_{2}M}{M^2-1}})$      \\
+    \midrule MPSQ     & $2Q(\sqrt{2\gamma_s}\sin(\pi/M))$                & $\frac{2}{\log_{2}M}Q(\sqrt{2\gamma_b\log_{2}M}\sin(\pi/M))$                \\
+    \midrule MQAM     & $4(1-1/\sqrt{M})Q(\sqrt{\frac{3\gamma_s}{M-1}})$ & $\frac{4(1-1/\sqrt{M})}{\log_{2}M}Q(\sqrt{\frac{3\gamma_b\log_{2}M}{M-1}})$ \\
+    \midrule FSK      & $(M-1)Q(\sqrt{\gamma_s/2})$                      & $\frac{(M-1)}{\log_{2} M}Q(\sqrt{\gamma_b \log_{2} M/2})$                   \\
     \bottomrule
@@
 \end{table}
+
+Where generally $P_b = P_s / \log{2}(M)$ for Gray coded constellations.
 
@@
 
+Another way to more easily calculate
+
+\section{Probability of Error for stochastic SNR}
+Since the SNR depends on the received power, which in turn depends on the
+affects of blocking, shadowing, and fading, the SNR is a random variable. It
+becomes useful then to define the average \gls{BER}.
+
+\begin{align}
+  \overline{P_b}  \triangleq \E{P_b} = \int_0^{\infty} P_{b}(\gamma) f_{\gamma}(\gamma) d\gamma
+\end{align}
+\begin{mdframed}
+
+  \subsection*{Average BPSK in Rayleigh Fading}
+  Express the average \gls{BER} of a BPSK signal in a Rayleigh fading channel in terms of the averge SNR $\overline{\gamma}$.
+  \\
+  \\
+  Answer: \\
+  For Rayleigh fading:
+  \begin{align}
+    f_{\gamma_s}(\gamma) = \frac{1}{\overline{\gamma}} e^{-\gamma/\overline{\gamma}}\nonumber
+  \end{align}
+  And for BPSK:
+  \begin{align}
+    P_b(\gamma_s) = 2Q(\sqrt{2\gamma_s}) \nonumber
+  \end{align}
+
+  So the \gls{BER} is:
+  \begin{align}
+    \overline{P_b} & = \int_0^{\infty} 2Q(\sqrt{2\gamma}) \frac{1}{\overline{\gamma}} e^{-\gamma/\overline{\gamma}} d\gamma \nonumber
+  \end{align}
+  with change of variables $x = \gamma/\overline{\gamma}$ such that $d \gamma/\overline{\gamma} = d x $
+
+  \begin{align}
+    \overline{P_b} & = \int_0^{\infty} 2Q(\sqrt{2x\overline{\gamma}}) e^{-x} dx \nonumber                                         \\
+                   & = \frac{2}{2\pi} \int_0^{\infty} e^{-x} \int_{\sqrt{2x\overline{\gamma}}}^{\infty} e^{-t^2/2} dtdx \nonumber
+  \end{align}
+  and switching the order of the integrals
+  \begin{align}
+    \overline{P_b} & = \frac{1}{\pi} \int_0^{\infty} \int_{0}^{t^2/(2\overline{\gamma})} e^{-t^2/2} e^{-x}  dxdt \nonumber \\
+                   & = \frac{1}{\pi} \int_0^{\infty} e^{-t^2/2}(1-e^{t^2/(2\overline{\gamma})}) dt \nonumber               \\
+                   & = \frac{1}{2} - \frac{1}{1\pi} \int_0^{\infty} e^{-t^2(1+1/\overline{\gamma})/2} dt \nonumber         \\
+                   & = \frac{1}{2} \left( 1-\sqrt{\frac{\overline{\gamma}}{1+\overline{\gamma}}}\right) \nonumber          \\
+                   & \approxeq \frac{1}{4\overline{\gamma}} \nonumber
+  \end{align}
+\end{mdframed}
+
+As seen in the example above it can become cumbersome to calculate the average bit error rate.
+To expedite the process we can use the moment generating functions of the channels and simplify the integral.
+Most constellation bit error probabilities can be approximated or precisely expressed as $$P_b(\gamma) = a Q( \sqrt{\beta\gamma})$$
+
+Using \cite{Craig258319}, we can express the $Q$-function as
+\begin{align}
+  Q(x) = \frac{1}{\pi}\int_{0}^{\pi/2} e^{-\frac{x^2}{2 \sin^2\phi}}d\phi
+\end{align}
+
 \printglossary[type=\acronymtype,title=Acronyms]
 
-% \bibliographystyle{ieeetran}
-% \bibliography{ref}
+\bibliographystyle{ieeetran}
+\bibliography{ref}
 \end{document}

### 2023-02-13 paper.tex
capacity

@@
 \newacronym{AWGN}{AWGN}{additive white Gaussian noise}
-
+\newacronym{rv}{r.v.}{random variable}
+\newacronym{CSI}{CSI}{channel side information}
+\newacronym{CDI}{CDI}{channel distribution information}
+\newacronym{WF}{WF}{waterfilling}
 \glsaddall
@@
 
-As seen in the example above it can become cumbersome to calculate the average bit error rate.
-To expedite the process we can use the moment generating functions of the channels and simplify the integral.
-Most constellation bit error probabilities can be approximated or precisely expressed as $$P_b(\gamma) = a Q( \sqrt{\beta\gamma})$$
+As seen in the example above it can become cumbersome to calculate the average
+bit error rate. To expedite the process we can use the moment generating
+functions of the channels and simplify the integral. Most constellation bit
+error probabilities can be approximated or precisely expressed as $$P_b(\gamma)
+  = a Q( \sqrt{\beta\gamma})$$
 
@@
 \end{align}
+TODO
 
+\chapter{Diversity}
+
+\chapter{Channel Capacity}
+
+\section{AWGN Channel Capacity, no fading}
+The capacity of a channel is a measure of the amount of information that can be
+transmitted over the channel, and is denoted with $C$. Another representation is
+with channel efficiency where the capacity is divided by the signal \gls{BW}
+denoted with B.
+
+\begin{align}
+  C   & = B log_{2}(1+\gamma) ~~ bps/ \text{complex channel}           \\
+  C/B & = log_{2}(1+\gamma) ~~ bps/Hz/ \text{complex channel\nonumber}
+\end{align}
+
+\begin{mdframed}
+
+  \subsection*{Shannon Capacity/ \gls{AWGN} Capacity derivation}
+  Derive the Shannon capacity using mutual information.
+  \\
+  \\
+  Answer: If an \gls{AWGN} channel $y = x + w$ where $w \sim \mathcal{N}(0,N_0)$
+  then the capacity is equal to the maximum amount of mutual information
+  $I(x,y)$ of the channel input and output.
+  \begin{align}
+    C/B & = \max_{x} I(x;y) = \max_{x} \left[h(y) - h(y|x) \right] \nonumber \\
+        & = \max_{x} \left[h(x+w) - h(x+w|x) \right] \nonumber               \\
+        & = \max_{x} \left[h(x+w) - h(w) \right] \nonumber
+  \end{align}
+
+  For a \gls{GRV} $w$, the entropy $h(w) = \frac{1}{2} \log_{2} (2 \pi e N_0)$.
+  Also, the distribution which maximizes differential entropy $h$ for random
+  variables of the same variance is Gaussian. This means maximizing $C$ can be
+  boiled down to maximizing $h(x+w)$, which happens when $x+w \sim
+    \mathcal{N}(\mu,\sigma^2)$. This is true only when $x \sim
+    \mathcal{N}(0,P_x)$.
+  \begin{align}
+    \therefore C/B & = \max_{x} I(x;y) = \frac{1}{2} \log_{2}(2\pi e (E_s + N_0)) - \frac{1}{2} \log_{2} (2 \pi e N_0) \nonumber \\
+                   & = \frac{1}{2} \log_{2}(1 + \frac{E_s}{N_0}) = \frac{1}{2} \log_{2}(1 + \gamma)
+  \end{align}
+  For $x$ a complex \gls{GRV} the capacity is doubled.
+\end{mdframed}
+
+\section{Flat Fading Channel Capacity}
+In this regime, the \gls{SNR} of the channel is a \gls{rv} due to fading.
+Since the noise is always modeled as Gaussian, what we model making the  \gls{SNR} of a channel a  \gls{rv} is the time-varying gain of a channel.
+The instantaneous received \gls{SNR} at time $i$ is a function of the average transmit power $\overline{P_x}$, the noise \gls{PSD}, signal \gls{BW} and instantaneous gain of the channel at time $i$ $g[i]$.
+
+The capacity of the channel, then, depends on the knowledge of the channel gain, called the \gls{CSI}. There are three choices regarding the \gls{CSI}:
+
+\begin{itemize}
+  \item \gls{CDI}: Both transmitter and receiver known the distribution of $g[i]$
+  \item Receiver \gls{CSI}: Only the receiver knows the value of $g[i]$, but both transmitter and receiver known the distribution of $g[i]$
+  \item Receiver \& transmitter \gls{CSI}:  Both transmitter and receiver known the distribution and the value of $g[i]$
+\end{itemize}
+
+\subsection{Channel Distribution Information}
+CDI is an open problem. It can become difficult to find the capacity-achieving
+distribution for the input. Literature exists on specific cases, such as the
+Rayleigh fading channel, in which the power gain is exponentially distrubuted
+and changes with each channel use at time $i$. An optimal solution exists, but
+the optimal distribution and capacity must be found numerically and lacks a
+closed-form solution.
+TODO
+\subsection{Receiver Channel Side Information}
+In this case the capacity is defined by the Shannon capacity.
+In practical systems the rate cannot be maintained constant as is expected in Shannon capacity, so a reduction from the Shannon capacity is expected.
+
+Shannon (ergodic) capacity can be computed by integrating over the distribution of the \gls{SNR}, which is known by both the transmitter and receiver.
+
+\begin{align}
+  C & = \E{B\log_{2}(1+\gamma)} =\int_0^{\infty}B \log_{2}(1+\gamma)p(\gamma)d\gamma
+\end{align}
+
+Since the realization of $\gamma$ is known only to the receiver, the ergotic capacity is constant, regardless of $\gamma$.
+A bound on the capacity can be roughly formed by using Jensen's inequality and switching the expectation and logarithm to achieve
+\begin{align}
+  C & \leq B\log_{2}(1+\E{\gamma}) = B\log_{2}(1+\overline{\gamma})
+\end{align}
+where $\overline{\gamma}$ is the average channel \gls{SNR}. This shows that for receiver \gls{CSI}, the achievable capacity is always less than the Shannon capacity.
+
+\subsection{Receiver \& Transmitter Channel Side Information}
+In this regime, the receiver can feed back information about the channel to the transmitter, and in turn the transmitter can take actions and adapt it's transmission strategy.
+
+There are two main knobs which the transmitter has access to in order to change it's strategy: power and rate.
+If we allow the transmit power to vary with some function of $\gamma$ subject to some average power constraint $\overline{P}$, we arrive at the power adaptation strategy.
+
+\subsection*{Water filling: Optimal Strategy}
+TODO
+Solving the resulting lagrangian from
+\begin{align}
+  C = \max_{P(\gamma)} \left[ \int_{0}^{\infty}B \log_2 (1 + \frac{P(\gamma)}{\overline{P}}\gamma)p(\gamma) d\gamma\right] \nonumber \\
+  \text{s.t. } \int P(\gamma)p(\gamma)d\gamma = \overline{P}
+\end{align}
+
+results in a \gls{WF} solution where
+\begin{align} \label{PA_policy}
+  \frac{P(\gamma)}{\overline{P}} = \left\{ \begin{array}{ll}
+                                             1/\gamma_0 - 1/\gamma & \gamma \geq \gamma_0 \\
+                                             0                     & \gamma < \gamma_0    \\
+                                           \end{array}
+  \right.
+\end{align}
+
+for the value of $\gamma_0$ that satisfies the power constraint from the lagrangian. The overall capacity in the receiver and transmitter \gls{CSI} is then:
+\begin{align}
+  C = \int_{\gamma_0}^{\infty} B \log_{2} (\frac{\gamma}{\gamma_0})p(\gamma)d\gamma
+\end{align}
+
+Since the power adaptation policy in (\ref{PA_policy}) has a dependence on the threshold, it can only be achieved with a time-varying data rate achieved for the instantaneous \gls{SNR}.
+For Rayleigh fading, this capacity can exceed the AWGN channel capacity for the same average SNR, different to the \gls{CSI} case where fading always indicated decreased capacity.
+
+For an arbitrary power adaptation policy, the channel capacity is
+\begin{align}
+  C = \int_{\gamma_0}^{\infty} B \log_{2} (1 + \frac{P(\gamma)}{\overline{P}}\gamma)p(\gamma)d\gamma
+\end{align}
+
+\subsection*{Discrete Rate adaptation}
+Since the \gls{WF} is not practically achievable, a discrete Rate adaptation method is described where the transmitter can choose the rate between discrete scemes (i.e. MQAM with varying M).
+TODO
+\subsection*{Power Adaptation at fixed rate}
+The power adaptation scheme employed here is:
+\begin{align} \label{PA_policy_fixedR}
+  \frac{P(\gamma)}{\overline{P}} = \left\{ \begin{array}{ll}
+                                             \gamma_t/\gamma & \gamma \geq \gamma_0 \\
+                                             0               & \gamma < \gamma_0    \\
+                                           \end{array}
+  \right.
+\end{align}
+for some target \gls{SNR} $\gamma_t$.
 \printglossary[type=\acronymtype,title=Acronyms]
